---
title: "EDA_for_8"
author: "Shuheng Jiang"
date: "2024-09-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

### Handle conflicts
```{r}
#| message: false
options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_ml.R?raw=true")
tidymodels_conflictRules()
# We also will need to resolve a new conflict using the following code
# Alternatively, John demonstrates code you can use when you load this library to prevent conflicts
conflictRules("Matrix", mask.ok = c("expand", "pack", "unpack"))
```


### Load required packages 
```{r}
#| message: false
#| warning: false
library(kableExtra, exclude = "group_rows") # for displaying formatted tables w/ kbl()
library(janitor, include.only = c("clean_names", "tabyl"))
library(cowplot, include.only = "plot_grid")
library(tidyverse) # for general data wrangling
library(tidymodels) # for modeling
library(kknn)
library(discrim, exclude = "smoothness")
library(xfun, include.only = "cache_rds")
library(cowplot, include.only = "plot_grid")
library(readxl)
library(dplyr)
library(lubridate)
library(skimr)
library(tidyselect)
library(tibble)
library(effectsize)
```


### Source function scripts
```{r}
#| message: false
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_plots.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
source("C:/Users/16089/OneDrive/Desktop/lab psych610/610710_functions.R")
```


### Specify other global settings
```{r}
#| include: false
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max=Inf)
```

### file path
```{r}
file_path <- "C:/Users/16089/OneDrive/Desktop/Greens lab/semester/raw_data_2"
```

## Cleaning EDA

### adjust the data columnnames 
#### for Questionnaire 
```{r}
data_1 <- read_csv(here::here(file_path, "Questionnaires_April 2, 2024_17.02_numbers.csv"), 
                   col_types = cols())
```

```{r}
data_2 <- read_csv(here::here(file_path, "Raven_April 2, 2024_16.11.csv"), 
                   col_types = cols())
```

```{r}
headers <- colnames(data_1)
d_num <- read_csv(here::here(file_path, "Questionnaires_April 2, 2024_17.02_numbers.csv"), skip = 2, col_names = TRUE)
colnames(d_num) <- headers
```

```{r}
headers <- colnames(data_2)
d_num_2 <- read_csv(here::here(file_path, "Raven_April 2, 2024_16.11.csv"), skip = 2, col_names = TRUE)
colnames(d_num_2) <- headers
```


### select the columns needed 
```{r}
d_num <- d_num %>%
  select(ParticipantID, EndDate, ResponseId, "SC0", "SC1", "SC2", "SC3", "SC4", "SC10", "SC11", "SC12", "SC13")
```

```{r}
d_num_2 <- d_num_2 %>% 
  select(ParticipantID, EndDate, ResponseId, Version, "SC0", Study, Q_TotalDuration)
```

## Cleaning of the Raven
```{r}

```

*Raven_April data*
*anything starts with "OA" under "study" column should be removed*
*Remove any row where ParticipantID is not "1XXX"*

*Keep Version, SC0,  column*
*ResponseID as participant_id_dataset*
*date*
*Q_TotalDuration*

```{r}
d_try <- d_num %>%
  rename(participant_id = ParticipantID,
         participant_id_dataset = ResponseId,
         date = EndDate)
```

```{r}
d_try_2 <- d_num_2 %>% 
  rename(participant_id = ParticipantID,
         participant_id_dataset = ResponseId,
         date = EndDate)
```

```{r}
d_try <- d_try |> 
  mutate(dataset_of_origin = "2021_JP_R56")
head(d_try)
```


```{r}
d_try <- d_try %>% select(participant_id, participant_id_dataset, date, dataset_of_origin, everything())
```

### Computing Scores
*For BFI score, divide by 2*
*Grit divide by 8*
*Conscientiousness divide by 10*
*Metacog divide by 30*
*Mindset divide by 20*

### BFI
```{r}
d_try$neuroticism_score <- d_try$SC3/2

d_try$conscientiousness_score <- d_try$SC2/2

d_try$openness_score <- d_try$SC4/2

d_try$extraversion_score <- d_try$SC0/2

d_try$agreeableness_score <- d_try$SC1/2
```

### Grit
```{r}
d_try$grit_score <- d_try$SC10/8
```

### Consciensciousness
```{r}
d_try$conscience_score <- d_try$SC11/10
```

### MetaCognition
```{r}
d_try$meta_cog <- d_try$SC12/30
```

### Mindset
```{r}
d_try$mindset_score <- d_try$SC13/20
```

### Computing Date
### adding time variables
#### year
```{r}
d_try$date <- as.POSIXct(d_try$date, format = "%Y-%m-%d %H:%M:%S")

# Extract the year and create a new variable
d_try$year <- year(d_try$date)

d_try <- d_try %>% select(participant_id, date, year, everything())
```


#### semester_id
```{r}
categorize_semester <- function(df) {
  # Ensure date is in the correct datetime format
  df$date <- as.POSIXct(df$date, format = "%Y-%m-%d %H:%M:%S")
  
  # Create the semester_id variable based on the date
  df <- df %>%
    mutate(semester_id = case_when(
      date >= as.POSIXct("2018-01-23") & date <= as.POSIXct("2018-05-11") ~ "Spring-2018",
      date >= as.POSIXct("2018-05-21") & date <= as.POSIXct("2018-06-15") ~ "Summer-2018-4-week",
      date >= as.POSIXct("2018-06-18") & date <= as.POSIXct("2018-08-10") ~ "Summer-2018-8-week",
      date >= as.POSIXct("2018-09-05") & date <= as.POSIXct("2018-12-20") ~ "Fall-2018",
      date >= as.POSIXct("2019-01-22") & date <= as.POSIXct("2019-05-10") ~ "Spring-2019",
      date >= as.POSIXct("2019-05-20") & date <= as.POSIXct("2019-06-14") ~ "Summer-2019-4-week",
      date >= as.POSIXct("2019-06-17") & date <= as.POSIXct("2019-08-09") ~ "Summer-2019-8-week",
      date >= as.POSIXct("2019-09-04") & date <= as.POSIXct("2019-12-19") ~ "Fall-2019",
      date >= as.POSIXct("2020-01-21") & date <= as.POSIXct("2020-05-08") ~ "Spring-2020",
      date >= as.POSIXct("2020-05-18") & date <= as.POSIXct("2020-6-12") ~ "Summer-2020-4-week",
      date >= as.POSIXct("2020-06-15") & date <= as.POSIXct("2020-08-07") ~ "Summer-2020-8-week",
      date >= as.POSIXct("2020-09-02") & date <= as.POSIXct("2020-12-18") ~ "Fall-2020",
      date >= as.POSIXct("2021-01-19") & date <= as.POSIXct("2021-05-07") ~ "Spring-2021",
      date >= as.POSIXct("2021-05-17") & date <= as.POSIXct("2021-06-11") ~ "Summer-2021-4-week",
      date >= as.POSIXct("2021-06-14") & date <= as.POSIXct("2021-08-06") ~ "Summer-2021-8-week",
      date >= as.POSIXct("2021-09-08") & date <= as.POSIXct("2021-12-19") ~ "Fall-2021",
      date >= as.POSIXct("2022-01-25") & date <= as.POSIXct("2022-05-13") ~ "Spring-2022",
      date >= as.POSIXct("2022-05-23") & date <= as.POSIXct("2022-06-17") ~ "Summer-2022-4-week",
      date >= as.POSIXct("2022-06-20") & date <= as.POSIXct("2022-08-12") ~ "Summer-2022-8-week",
      date >= as.POSIXct("2022-09-07") & date <= as.POSIXct("2022-12-22") ~ "Fall-2022",
      date >= as.POSIXct("2023-01-24") & date <= as.POSIXct("2023-05-12") ~ "Spring-2023",
      date >= as.POSIXct("2023-05-22") & date <= as.POSIXct("2022-06-16") ~ "Summer-2023-4-week",
      date >= as.POSIXct("2023-06-19") & date <= as.POSIXct("2023-08-11") ~ "Summer-2023-8-week",
      date >= as.POSIXct("2023-09-06") & date <= as.POSIXct("2023-12-21") ~ "Fall-2023",
      TRUE ~ NA_character_
    ))
  
  return(df)
}

# Apply the function to your data frame d_try
d_try <- categorize_semester(d_try)
```

```{r}
sum(is.na(d_try$semester_id))
```

#### academic_year 
```{r}
#(yyyy-xxxx)
add_academic_year <- function(dataframe) {
  dataframe$academic_year <- NA 
  dataframe$academic_year[dataframe$semester_id %in% c("Spring-2018", "Summer-2018-4-week", "Summer-2018-8-week")] <- "2017-2018"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2018", "Spring-2019", "Summer-2019-4-week", "Summer-2019-8-week")] <- "2018-2019"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2019", "Spring-2020", "Summer-2020-4-week", "Summer-2020-8-week")] <- "2019-2020"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2020", "Spring-2021", "Summer-2021-4-week", "Summer-2021-8-week")] <- "2020-2021"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2021", "Spring-2022", "Summer-2022-4-week", "Summer-2022-8-week")] <- "2021-2022"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2022", "Spring-2023", "Summer-2023-4-week", "Summer-2023-8-week")] <- "2022-2023"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2023")] <- "2023-2024"
  return(dataframe)
}
d_try <- add_academic_year(d_try)
```

```{r}
sum(is.na(d_try$academic_year))
```

#### semester
```{r}
#(“Fall”, “Spring”, “Summer-4” and “Summer-8”)
add_semester <- function(dataframe) {
  # Initialize the new variable with NA
  dataframe$semester <- NA
  
  # Define the pattern and replacement rules
  dataframe$semester[grep("Fall", dataframe$semester_id)] <- "Fall"
  dataframe$semester[grep("Spring", dataframe$semester_id)] <- "Spring"
  dataframe$semester[grep("Summer-4", dataframe$semester_id)] <- "Summer-4"
  dataframe$semester[grep("Summer-8", dataframe$semester_id)] <- "Summer-8"
  
  return(dataframe)
}

d_try <- add_semester(d_try)
```

#### first_day_of_the_semester
```{r}
add_first_day_of_the_semester <- function(dataframe) {
  # Initialize the new variable with NA
  dataframe$first_day_of_the_semester <- NA
  
  # Define the mapping of semester_id to first day of the semester
  first_day_mapping <- list(
    "Spring-2018" = "2018-01-23",
    "Summer-2018-4-week" = "2018-05-21",
    "Summer-2018-8-week" = "2018-06-18",
    "Fall-2018" = "2018-09-05",
    "Spring-2019" = "2019-01-22",
    "Summer-2019-4-week" = "2019-05-20",
    "Summer-2019-8-week" = "2019-06-17",
    "Fall-2019" = "2019-09-04", 
    "Spring-2020" = "2020-01-21", 
    "Summer-2020-4-week" = "2020-05-18", 
    "Summer-2020-8-week" = "2020-06-15", 
    "Fall-2020" = "2020-09-02",
    "Spring-2021" = "2021-01-19", 
    "Summer-2021-4-week" = "2021-05-17", 
    "Summer-2021-8-week" = "2021-06-14", 
    "Fall-2021" = "2021-09-08", 
    "Spring-2022" = "2022-01-25", 
    "Summer-2022-4-week" = "2022-05-23", 
    "Summer-2022-8-week" = "2022-06-20", 
    "Fall-2022" = "2022-09-07", 
    "Spring-2023" = "2023-01-24", 
    "Summer-2023-4-week" = "2023-05-22", 
    "Summer-2023-8-week" = "2023-06-19", 
    "Fall-2023" = "2023-09-06"
  )
  
  # Assign first day of the semester based on semester_id
  dataframe$first_day_of_the_semester <- sapply(dataframe$semester_id, function(x) first_day_mapping[[x]])
  
  return(dataframe)
}

d_try <- add_first_day_of_the_semester(d_try)

```


#### days_since_semester_start
```{r}
add_days_since_semester_start <- function(dataframe) {
  # Ensure date is in POSIXct format
  dataframe <- dataframe %>%
    mutate(date = as.POSIXct(date, format = "%Y-%m-%d %H:%M:%S"))

  # Ensure first_day_of_the_semester is in Date format
  dataframe <- dataframe %>%
    mutate(first_day_of_the_semester = as.Date(first_day_of_the_semester, format = "%Y-%m-%d"))
  
  # Calculate the days since the semester start
  dataframe <- dataframe %>%
    mutate(days_since_semester_start = round(as.numeric(difftime(date, first_day_of_the_semester, units = "days"))))
  
  return(dataframe)
}


d_try <- add_days_since_semester_start(d_try)
```


*clean composite above, compute the date as before for the EDA of this data*


