---
title: "Semester_Project"
author: "Shuheng Jiang"
date: "2024-04-26"
output: html_document
---
## Setup

### Handle conflicts
```{r}
#| message: false
options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_ml.R?raw=true")
tidymodels_conflictRules()
# We also will need to resolve a new conflict using the following code
# Alternatively, John demonstrates code you can use when you load this library to prevent conflicts
conflictRules("Matrix", mask.ok = c("expand", "pack", "unpack"))
```

### Load required packages 
```{r}
#| message: false
#| warning: false
library(kableExtra, exclude = "group_rows") # for displaying formatted tables w/ kbl()
library(janitor, include.only = c("clean_names", "tabyl"))
library(cowplot, include.only = "plot_grid")
library(tidyverse) # for general data wrangling
library(tidymodels) # for modeling
library(kknn)
library(discrim, exclude = "smoothness")
library(xfun, include.only = "cache_rds")
library(cowplot, include.only = "plot_grid")
library(readxl)
library(dplyr)
library(lubridate)
library(skimr)
library(tidyselect)
library(tibble)
library(effectsize)
```

### Source function scripts
```{r}
#| message: false
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_plots.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
source("C:/Users/16089/OneDrive/Desktop/lab psych610/610710_functions.R")
```

### Specify other global settings
```{r}
#| include: false
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max=Inf)
```

### file path
```{r}
file_path <- "C:/Users/16089/OneDrive/Desktop/Greens lab/semester/raw_data"
```


## Cleaning EDA

### adjust the data columnnames 
#### for 1718_num
```{r}
data_1 <- read_csv(here::here(file_path, "Generic post-study survey_1718_April 2, 2024_16.25_numbers.csv"), 
                   col_types = cols())
```


```{r message=FALSE, warning=FALSE}
headers <- colnames(data_1)
d_1718_num <- read_csv(here::here(file_path, "Generic post-study survey_1718_April 2, 2024_16.25_numbers.csv"), skip = 2, col_names = TRUE)
colnames(d_1718_num) <- headers
```


#### for other 3 datasets
```{r}
data_2 <- read_csv(here::here(file_path, "Generic post-study survey_2018-2019_April 2, 2024_16.26_numbers.csv"), 
                   col_types = cols())

data_3 <- read_csv(here::here(file_path, "Generic post-study survey_2019-2020_April 2, 2024_16.29_numbers.csv"), 
                   col_types = cols())

data_4 <- read_csv(here::here(file_path, "Generic post-study survey_2020-2021_April 2, 2024_16.33_numbers.csv"), 
                   col_types = cols())
```

#### Adjusting the column names
```{r}
headers_2 <- colnames(data_2)
d_1819_num <- read_csv(here::here(file_path, "Generic post-study survey_2018-2019_April 2, 2024_16.26_numbers.csv"), skip = 2, col_names = TRUE)
colnames(d_1819_num) <- headers_2

headers_3 <- colnames(data_3)
d_1920_num <- read_csv(here::here(file_path, "Generic post-study survey_2019-2020_April 2, 2024_16.29_numbers.csv"), skip = 2, col_names = TRUE)
colnames(d_1920_num) <- headers_3
```


```{r}
headers_4 <- colnames(data_4)
d_2021_num <- read_csv(here::here(file_path, "Generic post-study survey_2020-2021_April 2, 2024_16.33_numbers.csv"), skip = 2, col_names = TRUE)
colnames(d_2021_num) <- headers_4
```


### select the columns needed 
```{r}
d_1718_num_f <- d_1718_num %>%
  select(
    id, EndDate, ResponseId,
    starts_with("grit_"),
    starts_with("adhd_"),
    starts_with("neoPI_"),
    starts_with("growthMindSet_")
  )
```

```{r}
d_1819_num_f <- d_1819_num %>%
  select(
    id, EndDate, ResponseId,
    starts_with("grit_"),
    starts_with("adhd_"),
    starts_with("neoPI_"),
    starts_with("growthMindSet_"),
    ends_with("_r"), 
    ends_with("_a"),
    "3d5_A1B4C5"
  )
```

```{r}
names(d_1819_num_f)
```

```{r}
d_1920_num_f <- d_1920_num %>%
  select(
    id, EndDate, ResponseId,
    starts_with("grit_"),
    starts_with("adhd_"),
    starts_with("neoPI_"),
    starts_with("growthMindSet_"),
    ends_with("_r"), 
    ends_with("_a"),
    "3d5_A1B4C5"
  )

d_2021_num_f <- d_2021_num %>%
  select(
    id, EndDate, ResponseId,
    starts_with("grit_"),
    starts_with("adhd_"),
    starts_with("neoPI_"),
    starts_with("growthMindSet_"),
    ends_with("_r"), 
    ends_with("_a"),
    "3d5_A1B4C5"
  )
```

*notes: variables including grit, ADHD, intelligence(matrices tests), growth mindsets, neoPI(refer to the qualtrics about the corresponding column names, especially for the matrices tests), 5 variables in total. Change the column names for matrices tests. Maintain the id columns, and change id to participant_id. For 1718 dataset, we miss the data for matrices tests, and I found that for neoPI, the 1718 only has 50 columns.*


```{r}
# change the id to participant id
d_1718_num_f <- d_1718_num_f %>%
  rename(participant_id = id,
         participant_id_dataset = ResponseId,
         date = EndDate)
d_1819_num_f <- d_1819_num_f %>%
  rename(participant_id = id,
         participant_id_dataset = ResponseId,
         date = EndDate, 
         catch_question = grit_6,
         grit_6 = grit_7,
         grit_7 = grit_8,
         grit_8 = grit_9)
d_1920_num_f <- d_1920_num_f %>%
  rename(participant_id = id,
         participant_id_dataset = ResponseId,
         date = EndDate, 
         catch_question = grit_6,
         grit_6 = grit_7,
         grit_7 = grit_8,
         grit_8 = grit_9)
d_2021_num_f <- d_2021_num_f %>%
  rename(participant_id = id,
         participant_id_dataset = ResponseId,
         date = EndDate, 
         catch_question = grit_6,
         grit_6 = grit_7,
         grit_7 = grit_8,
         grit_8 = grit_9) 
```
*notes: because for the other three datasets, the grit_6 is a trivial question that tests whether the participants answering the questions with effort, grit_7, 8, 9 in other three equals to the grit_6, 7, 8 in 1718, so renaming the trivial question to be grit_9 for better computation of the grit score.* 

#### change the column names for neoPI for three datasets
```{r}
rename_columns <- function(column_names) {
  renamed_columns <- sapply(column_names, function(col) {
    # Check if the column name starts with "neoPI_"
    if (startsWith(col, "neoPI_")) {
      # Extract the number from the column name
      num <- as.numeric(sub("neoPI_", "", col))
      
      # Check if the extraction resulted in a numeric value
      if (!is.na(num)) {
        if (num >= 1 & num <= 18) {
          return(col)
        } else if (num == 19) {
          return("catch_question_1")
        } else if (num == 39) {
          return("catch_question_2")
        } else if (num >= 20 & num <= 38) {
          return(paste0("neoPI_", num - 1))
        } else if (num >= 40 & num <= 52) {
          return(paste0("neoPI_", num - 2))
        }
      } else {
        # If the number extraction fails, return the column name unchanged
        return(col)
      }
    } else {
      # If the column name does not start with "neoPI_", return it unchanged
      return(col)
    }
  })
  
  return(renamed_columns)
}

```
*Found the difference between other three datasets and 1718, that other three have two additional catech questions they are neoPI_19 and neoPI_39; with neoPI_19: On this question, please choose the response "Very Accurate" (the correct response is 5); neoPI_39: On this question, please choose the response "Moderately Inaccurate" (the correct response is 2); 1-very inaccurate, 2-moderately inaccurate, 3-Neither Inaccurate nor Accurate, 4-moderately accurate, 5-very accurate.*

```{r}
new_column_names <- rename_columns(colnames(d_1819_num_f))
colnames(d_1819_num_f) <- new_column_names
print(colnames(d_1819_num_f))
```

```{r}
new_column_names2 <- rename_columns(colnames(d_1920_num_f))
colnames(d_1920_num_f) <- new_column_names2
print(colnames(d_1920_num_f))
```

```{r}
new_column_names3 <- rename_columns(colnames(d_2021_num_f))
colnames(d_2021_num_f) <- new_column_names3
print(colnames(d_2021_num_f))
```

#### Handling the participant for catch questions 
```{r}
d_1819_num_f1 <- d_1819_num_f %>%
  filter(catch_question == 4 & catch_question_1 == 5 & catch_question_2 == 2)
d_1920_num_f1 <- d_1920_num_f %>% 
  filter(catch_question == 4 & catch_question_1 == 5 & catch_question_2 == 2)
d_2021_num_f1 <- d_2021_num_f %>% 
  filter(catch_question == 4 & catch_question_1 == 5 & catch_question_2 == 2)
```

```{r}
d_1819_num_wrong <- d_1819_num_f %>% 
  filter(catch_question != 4 | is.na(catch_question) | catch_question_1 != 5 | is.na(catch_question_1) | catch_question_2 != 2 | is.na(catch_question_2))
d_1920_num_wrong <- d_1920_num_f %>% 
  filter(catch_question != 4 | is.na(catch_question) | catch_question_1 != 5 | is.na(catch_question_1) | catch_question_2 != 2 | is.na(catch_question_2))
d_2021_num_wrong <- d_2021_num_f %>% 
  filter(catch_question != 4 | is.na(catch_question) | catch_question_1 != 5 | is.na(catch_question_1) | catch_question_2 != 2 | is.na(catch_question_2))
```

#### concatnating the wrong answer to catch question 
```{r}
d_wrong <- d_1819_num_wrong %>% full_join(d_1920_num_wrong) %>%
  full_join(d_2021_num_wrong) %>% as.data.frame()
```

### Concatnating four
```{r}
d_1718_num_f <- d_1718_num_f %>% mutate(source = "post-study-survey_2017-2018")
d_1819_num_f1 <- d_1819_num_f1 %>% mutate(source = "post-study-survey_2018-2019")
d_1920_num_f1 <- d_1920_num_f1 %>% mutate(source = "post-study-survey_2019-2020")
d_2021_num_f1 <- d_2021_num_f1 %>% mutate(source = "post-study-survey_2020-2021")
d_try <- d_1718_num_f %>% full_join(d_1819_num_f1) %>% full_join(d_1920_num_f1) %>%
  full_join(d_2021_num_f1) %>% as.data.frame()
```


### clean the column names and response 
```{r}
clean_class <- function(df){
  df |>
  clean_names("snake") |> 
  mutate(across(where(is.character), factor))|> 
  glimpse()
}
```

```{r}
d_wrong <- d_wrong |> clean_class()
```


```{r}
d_try <- d_try |> clean_class()
```

```{r}
d_try <- d_try |> 
  mutate(dataset_of_origin = "2017-2023_Generic_Post_Survey")
head(d_try)
```


## Computing the composite score for Grit 
```{r}
grit <- c("grit_1","grit_2","grit_3","grit_4","grit_5","grit_6","grit_7","grit_8" )
```

```{r}
d_try[,grit] %>% head()
```

```{r}
range(d_try[,grit], na.rm = TRUE)
```


### Computing the global grit score
```{r}
d_try$grit_overall = varScore(d_try, 
                   Forward= c("grit_1","grit_3","grit_5","grit_6"), 
                   Reverse= c("grit_2", "grit_4","grit_7","grit_8"), #reverse the forward and reverse, b/c of the grading issues
                   Range=c(1, 4),
                   Prorate=T, 
                   MaxMiss=0
)/8

```
*in our data, 1 means very like the person, 4 means not like the person, we want the final global score to be 4 very like the person and 1 not like the person, so we reverse the grading of the original score*

### Computing the score for Consistency of Interest 
```{r}
interest <- c("grit_1","grit_3","grit_5","grit_6")
d_try$grit_interest = varScore(d_try, 
                   interest, 
                   Range=c(1, 4),
                   Prorate=T, 
                   MaxMiss=0
)/4
```
*All items in Consistency of Interest are reverse coded items*

### Computing the score for Perseverance of Effort
```{r}
effort <- c("grit_2", "grit_4","grit_7","grit_8")
d_try$grit_effort = varScore_m(d_try, 
                   Reverse= effort, 
                   Range=c(1, 4),
                   Prorate=T, 
                   MaxMiss=0
)/4

```
*modified the varscore function to have the forward as empty*


## Compute the Composite score for ADHD

### Computing the binary score for all six items 
#### ADHD_1, 2, 3
```{r}
d_try$adhd_1_sig <- ifelse(d_try$adhd_1 == 3 | d_try$adhd_1 == 4 | d_try$adhd_1 == 5, 1, 0)
d_try$adhd_2_sig <- ifelse(d_try$adhd_2 == 3 | d_try$adhd_2 == 4 | d_try$adhd_2 == 5, 1, 0)
d_try$adhd_3_sig <- ifelse(d_try$adhd_3 == 3 | d_try$adhd_3 == 4 | d_try$adhd_3 == 5, 1, 0)
```
*threshold is above 3 for ADHD_1, 2, 3*

#### ADHD_4, 5, 6
```{r}
d_try$adhd_4_sig <- ifelse(d_try$adhd_4 == 4 | d_try$adhd_4 == 5, 1, 0)
d_try$adhd_5_sig <- ifelse(d_try$adhd_5 == 4 | d_try$adhd_5 == 5, 1, 0)
d_try$adhd_6_sig <- ifelse(d_try$adhd_6 == 4 | d_try$adhd_6 == 5, 1, 0)
```
*threshold is above 4 for ADHD_4, 5, 6*

### Computing the final diagnosis of ADHD
```{r}
adhd_sig <- c("adhd_1_sig", "adhd_2_sig", "adhd_3_sig", "adhd_4_sig", "adhd_5_sig", "adhd_6_sig")
d_try$adhd_total = varScore(d_try, 
                   adhd_sig,
                   Range=c(0, 1),
                   Prorate=T, 
                   MaxMiss=0
)
d_try$adhd_clinical <- ifelse(d_try$adhd_total == 4 | d_try$adhd_total == 5| d_try$adhd_total == 6, 1, 0)
```
*if 4 or more Qs are marked as clinically significant, we classify the person as positive in ADHD*

*for each measure, i.e. number of completed/number of total(complete rate), *

#### Cleaning the matrices test column names
```{r}
d_try <- d_try %>%
  rename(sandia_matrices_1 = x2d8_a1b3_a,
         ucmrt_1 = x02a395_r,
         sandia_matrices_2 = ld7_z_1_a,
         ucmrt_2 = x04a217_r,
         sandia_matrices_3 = x3d5_a1b4c5,
         ucmrt_3 = x09a364_r,
         sandia_matrices_4 = ld5_z_3_a,
         ucmrt_4  = x06a285_r,
         sandia_matrices_5 = x3d6_a1b3c5_a,
         ucmrt_5 = x10a398_r, 
         sandia_matrices_6 = ld5_x_1_a,
         ucmrt_6 = x11a347_r, 
         sandia_matrices_7 = x3d7_a1b2c4_a, 
         ucmrt_7 = x13a184_r,
         sandia_matrices_8 = ld5_z_4_a, 
         ucmrt_8 = x03a193_r,
         sandia_matrices_9 = ld6_y_1_a,
         ucmrt_9 = x08a147_r)
d_wrong <- d_wrong %>% 
    rename(sandia_matrices_1 = x2d8_a1b3_a,
         ucmrt_1 = x02a395_r,
         sandia_matrices_2 = ld7_z_1_a,
         ucmrt_2 = x04a217_r,
         sandia_matrices_3 = x3d5_a1b4c5,
         ucmrt_3 = x09a364_r,
         sandia_matrices_4 = ld5_z_3_a,
         ucmrt_4  = x06a285_r,
         sandia_matrices_5 = x3d6_a1b3c5_a,
         ucmrt_5 = x10a398_r, 
         sandia_matrices_6 = ld5_x_1_a,
         ucmrt_6 = x11a347_r, 
         sandia_matrices_7 = x3d7_a1b2c4_a, 
         ucmrt_7 = x13a184_r,
         sandia_matrices_8 = ld5_z_4_a, 
         ucmrt_8 = x03a193_r,
         sandia_matrices_9 = ld6_y_1_a,
         ucmrt_9 = x08a147_r)
```
*clean the column names of the matrices*
*2 questions missing in 1718, identify the 2 questions, then start computing the score, neoPI has 5 constructs, these 5 constructs are independent, we do not compute the global score for neoPI, but 5 scores for each construct*

#### skimming the separate dataframe for incorrect answers to catch questions
```{r}
d_wrong |> select("catch_question", "catch_question_1", "catch_question_2") |> 
  kbl(align = "r") |> 
  kable_styling(bootstrap_options = c("striped", "condensed")) |> 
  scroll_box(height = "500px", width = "100%")
```
*remove the participants who answer the catch question wrongly, and have them in a separate dataset*

#### changing participant_id for d_try
```{r}
d_try <- d_try %>%
  mutate(participant_id = sprintf('SEP_03_%04d', row_number()))
```
*assign new id to each participant, SE_0001, SE_0002, SE_0010, SE_0045, SE_0378, SE_0379, Etc....* 

### compute score for matrices 
#### assining score for each matries question
```{r}
d_try$sandia_1_correct <- ifelse(d_try$sandia_matrices_1 == 8, 1, 0)
d_try$sandia_2_correct <- ifelse(d_try$sandia_matrices_2 == 7, 1, 0)
d_try$sandia_3_correct <- ifelse(d_try$sandia_matrices_3 == 5, 1, 0)
d_try$sandia_4_correct <- ifelse(d_try$sandia_matrices_4 == 5, 1, 0)
d_try$sandia_5_correct <- ifelse(d_try$sandia_matrices_5 == 6, 1, 0)
d_try$sandia_6_correct <- ifelse(d_try$sandia_matrices_6 == 5, 1, 0)
d_try$sandia_7_correct <- ifelse(d_try$sandia_matrices_7 == 7, 1, 0)
d_try$sandia_8_correct <- ifelse(d_try$sandia_matrices_8 == 5, 1, 0)
d_try$sandia_9_correct <- ifelse(d_try$sandia_matrices_9 == 6, 1, 0)
```

```{r}
d_try$ucmrt_1_correct <- ifelse(d_try$ucmrt_1 == 5, 1, 0)
d_try$ucmrt_2_correct <- ifelse(d_try$ucmrt_2 == 7, 1, 0)
d_try$ucmrt_3_correct <- ifelse(d_try$ucmrt_3 == 6, 1, 0)
d_try$ucmrt_4_correct <- ifelse(d_try$ucmrt_4 == 2, 1, 0)
d_try$ucmrt_5_correct <- ifelse(d_try$ucmrt_5 == 1, 1, 0)
d_try$ucmrt_6_correct <- ifelse(d_try$ucmrt_6 == 8, 1, 0)
d_try$ucmrt_7_correct <- ifelse(d_try$ucmrt_7 == 2, 1, 0)
d_try$ucmrt_8_correct <- ifelse(d_try$ucmrt_8 == 6, 1, 0)
d_try$ucmrt_9_correct <- ifelse(d_try$ucmrt_9 == 5, 1, 0)
```

#### computing total score for sandia 
```{r}
sandia_matrices_score <- c("sandia_1_correct", "sandia_2_correct", "sandia_3_correct", "sandia_4_correct", "sandia_5_correct", "sandia_6_correct", "sandia_7_correct", "sandia_8_correct", "sandia_9_correct")
d_try$sandia_matrices_hit_rate = varScore(d_try, 
                   sandia_matrices_score,
                   Range=c(0, 1),
                   Prorate=T, 
                   MaxMiss=0
)/9
```

#### computing total score for ucmrt
```{r}
ucmrt_score <- c("ucmrt_1_correct", "ucmrt_2_correct", "ucmrt_3_correct", "ucmrt_4_correct", "ucmrt_5_correct", "ucmrt_6_correct", "ucmrt_7_correct", "ucmrt_8_correct", "ucmrt_9_correct")
d_try$ucmrt_hit_rate = varScore(d_try, 
                   ucmrt_score,
                   Range=c(0, 1),
                   Prorate=T, 
                   MaxMiss=0
)/9
```

#### computing total score for matrices test
```{r}
matrices_sccore <- c(ucmrt_score, sandia_matrices_score)
d_try$matrices_hit_rate = varScore(d_try, 
                   matrices_sccore,
                   Range=c(0, 1),
                   Prorate=T, 
                   MaxMiss=0
)/18
```
*compute the score for the matrices, compute the average score*

### Comuting score for NeoPI
```{r}
neuroticism <- c("neo_pi_1", "neo_pi_9", "neo_pi_10", "neo_pi_15", "neo_pi_22", "neo_pi_27", "neo_pi_34", "neo_pi_39", "neo_pi_43","neo_pi_47")
conscientiousness <- c("neo_pi_2", "neo_pi_7", "neo_pi_11", "neo_pi_16", "neo_pi_28", "neo_pi_29", "neo_pi_30", "neo_pi_35", "neo_pi_44","neo_pi_50")
openness <- c("neo_pi_4", "neo_pi_6", "neo_pi_12", "neo_pi_18", "neo_pi_19", "neo_pi_23", "neo_pi_25", "neo_pi_32", "neo_pi_37","neo_pi_41")
extraversion <- c("neo_pi_5", "neo_pi_13", "neo_pi_14", "neo_pi_20", "neo_pi_24", "neo_pi_33", "neo_pi_38", "neo_pi_42", "neo_pi_46","neo_pi_49")
agreeableness <- c("neo_pi_3", "neo_pi_8", "neo_pi_17", "neo_pi_21", "neo_pi_26", "neo_pi_31", "neo_pi_36", "neo_pi_40", "neo_pi_45","neo_pi_48")
```

```{r}
d_try$neuroticism_score = varScore_m(d_try, 
                   Forward = c("neo_pi_1", "neo_pi_15", "neo_pi_27", "neo_pi_39","neo_pi_47"),
                   Reverse= c("neo_pi_9", "neo_pi_10", "neo_pi_22", "neo_pi_34", "neo_pi_43"),
                   Range=c(1, 5),
                   Prorate=T, 
                   MaxMiss=0
)/10
d_try$conscientiousness_score = varScore_m(d_try, 
                   Forward = c("neo_pi_7", "neo_pi_16", "neo_pi_28", "neo_pi_29","neo_pi_44"),
                   Reverse= c("neo_pi_2", "neo_pi_11", "neo_pi_30", "neo_pi_35", "neo_pi_50"),
                   Range=c(1, 5),
                   Prorate=T, 
                   MaxMiss=0
)/10
d_try$openness_score = varScore_m(d_try, 
                   Forward = c("neo_pi_4", "neo_pi_12", "neo_pi_18", "neo_pi_25","neo_pi_41"),
                   Reverse= c("neo_pi_6", "neo_pi_19", "neo_pi_23", "neo_pi_32", "neo_pi_37"),
                   Range=c(1, 5),
                   Prorate=T, 
                   MaxMiss=0
)/10
d_try$extraversion_score = varScore_m(d_try, 
                   Forward = c("neo_pi_5", "neo_pi_13", "neo_pi_33", "neo_pi_38","neo_pi_49"),
                   Reverse= c("neo_pi_14", "neo_pi_20", "neo_pi_24", "neo_pi_42","neo_pi_46"),
                   Range=c(1, 5),
                   Prorate=T, 
                   MaxMiss=0
)/10
d_try$agreeableness_score = varScore_m(d_try, 
                   Forward = c("neo_pi_3", "neo_pi_17", "neo_pi_26", "neo_pi_36","neo_pi_48"),
                   Reverse= c("neo_pi_8", "neo_pi_21", "neo_pi_31", "neo_pi_40", "neo_pi_45"),
                   Range=c(1, 5),
                   Prorate=T, 
                   MaxMiss=0
)/10
```
*compute the score for the neoPI, higher score means higher accuracy of description*

### Computing Score for Growth Mindset
By checking the raw data, questions order is consistent across all 5 dataframes; 
substract 5 from all the score, Q2, Q4, Q6, Q8 are reverse coded
```{r}
d_try$growth_mind_set_1 <- d_try$growth_mind_set_1 - 5 
d_try$growth_mind_set_2 <- d_try$growth_mind_set_2 - 5 
d_try$growth_mind_set_3 <- d_try$growth_mind_set_3 - 5 
d_try$growth_mind_set_4 <- d_try$growth_mind_set_4 - 5 
d_try$growth_mind_set_5 <- d_try$growth_mind_set_5 - 5 
d_try$growth_mind_set_6 <- d_try$growth_mind_set_6 - 5 
d_try$growth_mind_set_7 <- d_try$growth_mind_set_7 - 5 
d_try$growth_mind_set_8 <- d_try$growth_mind_set_8 - 5 

psych::describe(d_try$growth_mind_set_1)
d_try$growth_mindset_score = varScore_m(d_try, 
                   Forward = c("growth_mind_set_1", "growth_mind_set_3", "growth_mind_set_5", "growth_mind_set_7"),
                   Reverse= c("growth_mind_set_2", "growth_mind_set_4", "growth_mind_set_6", "growth_mind_set_8"),
                   Range=c(1, 7),
                   Prorate=T, 
                   MaxMiss=0
)/8

```
Finished Computing Composite Score!  

### Printing the complete rate
*Count the number of participant for each measures with complete data(no NAs), count the percentage of participants that have complete data*
```{r}
d_try %>%
  skim_some(grit_overall, matrices_hit_rate, ucmrt_hit_rate, sandia_matrices_hit_rate, neuroticism_score, extraversion_score, conscientiousness_score, openness_score, agreeableness_score, adhd_clinical, growth_mindset_score) %>% 
  select(skim_variable, n_missing, complete_rate)
```

*n_complete: nrows - n_missing*
*How many people that answer the catch questions wrongly, and the percentage for each semester, and in total number of people who answer the catch questions wrongly at each point of the semester* 

```{r}
d_try %>%
  skim_some_m(grit_overall, matrices_hit_rate, ucmrt_hit_rate, sandia_matrices_hit_rate, neuroticism_score, extraversion_score, conscientiousness_score, openness_score, agreeableness_score, adhd_clinical, growth_mindset_score) %>% 
  select(skim_variable, n_complete, complete_rate)
```
*n_complete, n_missing, n_total, complete_rate; put them together in the same table, big5_neuroticism, big5_openness..*

#### printing something for d_wrong
```{r}
# having no idea about demographic data for d_wrong 
```

### adding time variables
#### year
```{r}
d_try$date <- as.POSIXct(d_try$date, format = "%Y-%m-%d %H:%M:%S")

# Extract the year and create a new variable
d_try$year <- year(d_try$date)

d_try <- d_try %>% select(participant_id, date, year, everything())
```

#### semester_id
```{r}
categorize_semester <- function(df) {
  # Ensure date is in the correct datetime format
  df$date <- as.POSIXct(df$date, format = "%Y-%m-%d %H:%M:%S")
  
  # Create the semester_id variable based on the date
  df <- df %>%
    mutate(semester_id = case_when(
      date >= as.POSIXct("2018-01-23") & date <= as.POSIXct("2018-05-11") ~ "Spring-2018",
      date >= as.POSIXct("2018-05-21") & date <= as.POSIXct("2018-06-15") ~ "Summer-2018-4-week",
      date >= as.POSIXct("2018-06-18") & date <= as.POSIXct("2018-08-10") ~ "Summer-2018-8-week",
      date >= as.POSIXct("2018-09-05") & date <= as.POSIXct("2018-12-20") ~ "Fall-2018",
      date >= as.POSIXct("2019-01-22") & date <= as.POSIXct("2019-05-10") ~ "Spring-2019",
      date >= as.POSIXct("2019-05-20") & date <= as.POSIXct("2019-06-14") ~ "Summer-2019-4-week",
      date >= as.POSIXct("2019-06-17") & date <= as.POSIXct("2019-08-09") ~ "Summer-2019-8-week",
      date >= as.POSIXct("2019-09-04") & date <= as.POSIXct("2019-12-19") ~ "Fall-2019",
      date >= as.POSIXct("2020-01-21") & date <= as.POSIXct("2020-05-08") ~ "Spring-2020",
      date >= as.POSIXct("2020-05-18") & date <= as.POSIXct("2020-6-12") ~ "Summer-2020-4-week",
      date >= as.POSIXct("2020-06-15") & date <= as.POSIXct("2020-08-07") ~ "Summer-2020-8-week",
      date >= as.POSIXct("2020-09-02") & date <= as.POSIXct("2020-12-18") ~ "Fall-2020",
      date >= as.POSIXct("2021-01-19") & date <= as.POSIXct("2021-05-07") ~ "Spring-2021",
      date >= as.POSIXct("2021-05-17") & date <= as.POSIXct("2021-06-11") ~ "Summer-2021-4-week",
      date >= as.POSIXct("2021-06-14") & date <= as.POSIXct("2021-08-06") ~ "Summer-2021-8-week",
      date >= as.POSIXct("2021-09-08") & date <= as.POSIXct("2021-12-19") ~ "Fall-2021",
      date >= as.POSIXct("2022-01-25") & date <= as.POSIXct("2022-05-13") ~ "Spring-2022",
      date >= as.POSIXct("2022-05-23") & date <= as.POSIXct("2022-06-17") ~ "Summer-2022-4-week",
      date >= as.POSIXct("2022-06-20") & date <= as.POSIXct("2022-08-12") ~ "Summer-2022-8-week",
      date >= as.POSIXct("2022-09-07") & date <= as.POSIXct("2022-12-22") ~ "Fall-2022",
      date >= as.POSIXct("2023-01-24") & date <= as.POSIXct("2023-05-12") ~ "Spring-2023",
      date >= as.POSIXct("2023-05-22") & date <= as.POSIXct("2023-06-16") ~ "Summer-2023-4-week",
      date >= as.POSIXct("2023-06-19") & date <= as.POSIXct("2023-08-11") ~ "Summer-2023-8-week",
      date >= as.POSIXct("2023-09-06") & date <= as.POSIXct("2023-12-21") ~ "Fall-2023",
      TRUE ~ NA_character_
    ))
  
  return(df)
}

# Apply the function to your data frame d_try
d_try <- categorize_semester(d_try)
```

```{r}
sum(is.na(d_try$semester_id))
```

#### academic_year 
```{r}
#(yyyy-xxxx)
add_academic_year <- function(dataframe) {
  dataframe$academic_year <- NA 
  dataframe$academic_year[dataframe$semester_id %in% c("Spring-2018", "Summer-2018-4-week", "Summer-2018-8-week")] <- "2017-2018"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2018", "Spring-2019", "Summer-2019-4-week", "Summer-2019-8-week")] <- "2018-2019"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2019", "Spring-2020", "Summer-2020-4-week", "Summer-2020-8-week")] <- "2019-2020"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2020", "Spring-2021", "Summer-2021-4-week", "Summer-2021-8-week")] <- "2020-2021"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2021", "Spring-2022", "Summer-2022-4-week", "Summer-2022-8-week")] <- "2021-2022"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2022", "Spring-2023", "Summer-2023-4-week", "Summer-2023-8-week")] <- "2022-2023"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2023")] <- "2023-2024"
  return(dataframe)
}
d_try <- add_academic_year(d_try)
```

```{r}
sum(is.na(d_try$academic_year))
```

#### semester
```{r}
#(“Fall”, “Spring”, “Summer-4” and “Summer-8”)
add_semester <- function(dataframe) {
  # Initialize the new variable with NA
  dataframe$semester <- NA
  
  # Define the pattern and replacement rules
  dataframe$semester[grep("Fall", dataframe$semester_id)] <- "Fall"
  dataframe$semester[grep("Spring", dataframe$semester_id)] <- "Spring"
  dataframe$semester[grep("Summer-4", dataframe$semester_id)] <- "Summer-4"
  dataframe$semester[grep("Summer-8", dataframe$semester_id)] <- "Summer-8"
  
  return(dataframe)
}

d_try <- add_semester(d_try)
```

#### first_day_of_the_semester
```{r}
add_first_day_of_the_semester <- function(dataframe) {
  # Initialize the new variable with NA
  dataframe$first_day_of_the_semester <- NA
  
  # Define the mapping of semester_id to first day of the semester
  first_day_mapping <- list(
    "Spring-2018" = "2018-01-23",
    "Summer-2018-4-week" = "2018-05-21",
    "Summer-2018-8-week" = "2018-06-18",
    "Fall-2018" = "2018-09-05",
    "Spring-2019" = "2019-01-22",
    "Summer-2019-4-week" = "2019-05-20",
    "Summer-2019-8-week" = "2019-06-17",
    "Fall-2019" = "2019-09-04", 
    "Spring-2020" = "2020-01-21", 
    "Summer-2020-4-week" = "2020-05-18", 
    "Summer-2020-8-week" = "2020-06-15", 
    "Fall-2020" = "2020-09-02",
    "Spring-2021" = "2021-01-19", 
    "Summer-2021-4-week" = "2021-05-17", 
    "Summer-2021-8-week" = "2021-06-14", 
    "Fall-2021" = "2021-09-08", 
    "Spring-2022" = "2022-01-25", 
    "Summer-2022-4-week" = "2022-05-23", 
    "Summer-2022-8-week" = "2022-06-20", 
    "Fall-2022" = "2022-09-07", 
    "Spring-2023" = "2023-01-24", 
    "Summer-2023-4-week" = "2023-05-22", 
    "Summer-2023-8-week" = "2023-06-19", 
    "Fall-2023" = "2023-09-06"
  )
  
  # Assign first day of the semester based on semester_id
  dataframe$first_day_of_the_semester <- sapply(dataframe$semester_id, function(x) first_day_mapping[[x]])
  
  return(dataframe)
}

d_try <- add_first_day_of_the_semester(d_try)

```


#### days_since_semester_start
```{r}
add_days_since_semester_start <- function(dataframe) {
  # Ensure date is in POSIXct format
  dataframe <- dataframe %>%
    mutate(date = as.POSIXct(date, format = "%Y-%m-%d %H:%M:%S"))

  # Ensure first_day_of_the_semester is in Date format
  dataframe <- dataframe %>%
    mutate(first_day_of_the_semester = as.Date(first_day_of_the_semester, format = "%Y-%m-%d"))
  
  # Calculate the days since the semester start
  dataframe <- dataframe %>%
    mutate(days_since_semester_start = round(as.numeric(difftime(date, first_day_of_the_semester, units = "days"))))
  
  return(dataframe)
}


d_try <- add_days_since_semester_start(d_try)

d_try <- d_try %>% select(participant_id, participant_id_dataset, semester_id, date, year, academic_year, semester, days_since_semester_start, first_day_of_the_semester, dataset_of_origin, source, grit_overall, matrices_hit_rate, ucmrt_hit_rate, sandia_matrices_hit_rate, neuroticism_score, extraversion_score, conscientiousness_score, openness_score, agreeableness_score, adhd_clinical,adhd_total, growth_mindset_score, everything())
```
*Notes: 1) the end date is end of class date, but many surveys were taken on the study day, include the exam day; 2) Could not be sure whether the students had taken summer classes or not, probably who had response during the summer must had been taking summer classes;(Assume participant who had response during the 4-week session enrolled in 4-week session and who responded during 8-week session enrolled in 8-week session);  3)making the new variable: just need to record the number of days since the beginning of the semester* 


*Thinking about the questions we can answer with the datasets:1)is there difference from the beginning, middle, and end of the semester, the number of days, the day in the semester the participant came in to do the survey, make the variables, academic year:2017-2018, 2018-2019, 2019-2020, 2020-2021, 2021-2022, 2022-2023, 2023-2024; semester: fall, spring, summer; point of the semester:the number of days since the beginning of the semester; the first day of the semester for each of the semester*


### Visual Exploration
#### Quick plot, histogram, distribution 
```{r}
ggplot(data = d_try) + 
  geom_bar(aes(x= grit_overall))
```

```{r}
# Might be something wrong with the culculation 
# hist(d_try$grit_interest, breaks = 10)
ggplot(data = d_try) + 
  geom_bar(aes(x= grit_interest))
```

```{r}
# Might be something wrong with the calculation 
ggplot(data = d_try) + 
  geom_bar(aes(x= grit_effort))
```

```{r}
ggplot(data = d_try) + 
  geom_bar(aes(x= adhd_total))
```

```{r}
ggplot(data = d_try) + 
  geom_bar(aes(x= adhd_clinical))
```

```{r}
ggplot(data = d_try) + 
  geom_bar(aes(x= matrices_hit_rate))
```

```{r}
ggplot(data = d_try) + 
  geom_bar(aes(x= sandia_matrices_hit_rate))
```
```{r}
ggplot(data = d_try) + 
  geom_bar(aes(x= ucmrt_hit_rate))
```

```{r}
ggplot(data = d_try) + 
  geom_bar(aes(x= neuroticism_score))
```

*delete start_date, recorded_date, replace date with date*
*Add a column named "dataset_of_origin"*
*Include "ResponseId" column in the raw data, change name into "participant_id_dataset"* 

### Output as csv
```{r}
d_try |> write_csv(here::here(file_path, "Generic_Post_Survey_Combined.csv"))
d_wrong |> write_csv(here::here(file_path, 
"Generic_Post_Survey_Combined_Excluded.csv"))
```
