---
title: "Visualizations_of_three_datasets"
author: "Shuheng Jiang"
date: "2024-10-11"
output: html_document
---
## Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Handle conflicts
```{r}
#| message: false
options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_ml.R?raw=true")
tidymodels_conflictRules()
# We also will need to resolve a new conflict using the following code
# Alternatively, John demonstrates code you can use when you load this library to prevent conflicts
conflictRules("Matrix", mask.ok = c("expand", "pack", "unpack"))
```

### Load required packages 
```{r}
#| message: false
#| warning: false
library(kableExtra, exclude = "group_rows") # for displaying formatted tables w/ kbl()
library(janitor, include.only = c("clean_names", "tabyl"))
library(cowplot, include.only = "plot_grid")
library(tidyverse) # for general data wrangling
library(tidymodels) # for modeling
library(kknn)
library(discrim, exclude = "smoothness")
library(xfun, include.only = "cache_rds")
library(cowplot, include.only = "plot_grid")
library(readxl)
library(dplyr)
library(lubridate)
library(skimr)
library(tidyselect)
library(tibble)
library(effectsize)
library(ggpubr)
library(latltools)
```


### Source function scripts
```{r}
#| message: false
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_plots.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
source("C:/Users/16089/OneDrive/Desktop/lab psych610/610710_functions.R")
```

### Specify other global settings
```{r}
#| include: false
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max=Inf)
```

### file path
```{r}
file_path <- "C:/Users/16089/OneDrive/Desktop/Capstone_Project/semester_effect/raw_data_3"
cleaned_path <- "C:/Users/16089/OneDrive/Desktop/Capstone_Project/semester_effect/cleaned_data_3"
```

### reading in data
```{r}
d_personality <- read_csv(here::here(file_path, "2021_EY_FYP.csv"), 
                   col_types = cols())
```

```{r}
d_torr <- read_csv(here::here(file_path, "2022_LA_Patterns_FYP.csv"), 
                   col_types = cols())
```

```{r}
d_matrices <- read_csv(here::here(file_path, "2021_AC_WMGf.csv"), 
                   col_types = cols())
```

### Cleaning the data

#### Cleaning the outlier
```{r}
d_personality<- d_personality[!is.na(d_personality$days_since_semester_start), ]
range(d_personality$days_since_semester_start)
```

```{r}
d_torr <- d_torr[!is.na(d_torr$days_since_semester_start), ]
range(d_torr$days_since_semester_start)
```


```{r}
hist(d_personality$days_since_semester_start)
# Do not know why there are no beginning of the semester data
```



```{r}
head(d_personality[order(d_personality$days_since_semester_start), ], n = 50) 
```


```{r}
hist(d_torr$days_since_semester_start)
```

remove outliers
```{r}
tail(d_torr[order(d_torr$days_since_semester_start), ], n = 50) # 10 highest values
```


```{r}
hist(d_matrices$days_since_semester_start)
```

```{r}
d_torr$first_day_of_the_semester[d_torr$semester_id == "summer-8_2020" & is.na(d_torr$first_day_of_the_semester)] <- "2020-06-20"
```

```{r}
d_torr$date <- as.Date(d_torr$date)
d_torr$first_day_of_the_semester <- as.Date(d_torr$first_day_of_the_semester)

# Calculate the difference and fill in the 'days_since_semester_start' column
d_torr <- d_torr |>
mutate(days_since_semester_start = round(as.numeric(difftime(date, first_day_of_the_semester, units = "days"))))
```

```{r}
head(d_torr[order(d_torr$days_since_semester_start), ], n = 50) # 10 highest values
```

```{r}
tail(d_torr[order(d_torr$days_since_semester_start), ], n = 50) # 10 highest values
```


```{r}
hist(d_torr$days_since_semester_start)
```

```{r}
d_torr <- d_torr[!is.na(d_torr$days_since_semester_start), ]
range(d_torr$days_since_semester_start)
```

```{r}
d_matrices <- d_matrices[!is.na(d_matrices$days_since_semester_start), ]
range(d_matrices$days_since_semester_start)
```

### for d_matrices Cleaning the Missing composite score
```{r}
names(d_matrices)
```

```{r}
d_matrices$sandia_1_correct <- ifelse(d_matrices$Q34 == 8, 1, 0)
d_matrices$sandia_2_correct <- ifelse(d_matrices$Q36 == 5, 1, 0)
d_matrices$sandia_3_correct <- ifelse(d_matrices$Q38 == 7, 1, 0)
d_matrices$sandia_4_correct <- ifelse(d_matrices$Q40 == 5, 1, 0)
d_matrices$sandia_5_correct <- ifelse(d_matrices$Q42 == 7, 1, 0)
d_matrices$sandia_6_correct <- ifelse(d_matrices$Q44 == 5, 1, 0)
d_matrices$sandia_7_correct <- ifelse(d_matrices$Q46 == 6, 1, 0)
d_matrices$sandia_8_correct <- ifelse(d_matrices$Q48 == 7, 1, 0)
d_matrices$sandia_9_correct <- ifelse(d_matrices$Q50 == 5, 1, 0)
d_matrices$sandia_10_correct <- ifelse(d_matrices$Q52 ==2, 1, 0)
d_matrices$sandia_11_correct <- ifelse(d_matrices$Q54 == 7, 1, 0)
d_matrices$sandia_12_correct <- ifelse(d_matrices$Q56 == 2, 1, 0)
d_matrices$sandia_13_correct <- ifelse(d_matrices$Q58 == 5, 1, 0)
d_matrices$sandia_14_correct <- ifelse(d_matrices$Q60 == 6, 1, 0)
```

```{r}
sandia_matrices_score <- c("sandia_1_correct", "sandia_2_correct", "sandia_3_correct", "sandia_4_correct", "sandia_5_correct", "sandia_6_correct", "sandia_7_correct", "sandia_8_correct", "sandia_9_correct", "sandia_10_correct", "sandia_11_correct", "sandia_12_correct", "sandia_13_correct", "sandia_14_correct")
d_matrices$sandia_matrices_hit_rate = varScore(d_matrices, 
                   sandia_matrices_score,
                   Range=c(0, 1),
                   Prorate=T, 
                   MaxMiss=0
)/14
```

```{r}
d_matrices$ucmrt_1_correct <- ifelse(d_matrices$Q91 == 5, 1, 0)
d_matrices$ucmrt_2_correct <- ifelse(d_matrices$Q93 == 8, 1, 0)
d_matrices$ucmrt_3_correct <- ifelse(d_matrices$Q95 == 3, 1, 0)
d_matrices$ucmrt_4_correct <- ifelse(d_matrices$Q97 == 6, 1, 0)
d_matrices$ucmrt_5_correct <- ifelse(d_matrices$Q99 == 8, 1, 0)
d_matrices$ucmrt_6_correct <- ifelse(d_matrices$Q101 == 8, 1, 0)
d_matrices$ucmrt_7_correct <- ifelse(d_matrices$Q103 == 6, 1, 0)
d_matrices$ucmrt_8_correct <- ifelse(d_matrices$Q105 == 7, 1, 0)
d_matrices$ucmrt_9_correct <- ifelse(d_matrices$Q107 == 5, 1, 0)
d_matrices$ucmrt_10_correct <- ifelse(d_matrices$Q109 == 6, 1, 0)
d_matrices$ucmrt_11_correct <- ifelse(d_matrices$Q111 == 8, 1, 0)
d_matrices$ucmrt_12_correct <- ifelse(d_matrices$Q113 == 6, 1, 0)
d_matrices$ucmrt_13_correct <- ifelse(d_matrices$Q115 == 7, 1, 0)
d_matrices$ucmrt_14_correct <- ifelse(d_matrices$Q117 == 1, 1, 0)
d_matrices$ucmrt_15_correct <- ifelse(d_matrices$Q119 == 1, 1, 0)
d_matrices$ucmrt_16_correct <- ifelse(d_matrices$Q121 == 1, 1, 0)
d_matrices$ucmrt_17_correct <- ifelse(d_matrices$Q123 == 5, 1, 0)
d_matrices$ucmrt_18_correct <- ifelse(d_matrices$Q125 == 3, 1, 0)
```

```{r}
ucmrt_score <- c("ucmrt_1_correct", "ucmrt_2_correct", "ucmrt_3_correct", "ucmrt_4_correct", "ucmrt_5_correct", "ucmrt_6_correct", "ucmrt_7_correct", "ucmrt_8_correct", "ucmrt_9_correct", "ucmrt_10_correct", "ucmrt_11_correct", "ucmrt_12_correct", "ucmrt_13_correct", "ucmrt_14_correct", "ucmrt_15_correct", "ucmrt_16_correct", "ucmrt_17_correct", "ucmrt_18_correct")
d_matrices$ucmrt_hit_rate = varScore(d_matrices, 
                   ucmrt_score,
                   Range=c(0, 1),
                   Prorate=T, 
                   MaxMiss=0
)/18
```

```{r}
matrices_sccore <- c(ucmrt_score, sandia_matrices_score)
d_matrices$matrices_hit_rate = varScore(d_matrices, 
                   matrices_sccore,
                   Range=c(0, 1),
                   Prorate=T, 
                   MaxMiss=0
)/32
```

```{r}
d_matrices <- d_matrices |>
  relocate(ucmrt_hit_rate,sandia_matrices_hit_rate,matrices_hit_rate, .after = dataset_of_origin)
```

Finished Cleaning d_matrices

### for d_torr
```{r}
d_torr <- d_torr |>
  relocate(TORR_Set1_Score,TORR_Set2_Score, ATMI_Score,WJ_Score, .after = dataset_of_origin) |>
  clean_names()
```

### for d_personaity Standadize the personality as post-survey
Not sure how to standardize the personality sore in this case, because they are different scale.
```{r}
d_personality <- d_personality |>
  relocate(extraversion, agreeableness, conscientiousness, neuroticism, openness, .after = dataset_of_origin) |>
  clean_names()
```

## Save CSV
```{r}
d_personality|> write_csv(here::here(cleaned_path, "Clean_2021_EY_FYP_Personality.csv"))
d_torr |> write_csv(here::here(cleaned_path, "Clean_2022_LA_Patterns_FYP_torr.csv"))
d_matrices |> write_csv(here::here(cleaned_path, "Clean_AC_2021_wmgf_matrices.csv"))
```

