---
title: "EDA_for_8"
author: "Shuheng Jiang"
date: "2024-09-12"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

### Handle conflicts
```{r}
#| message: false
options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_ml.R?raw=true")
tidymodels_conflictRules()
# We also will need to resolve a new conflict using the following code
# Alternatively, John demonstrates code you can use when you load this library to prevent conflicts
conflictRules("Matrix", mask.ok = c("expand", "pack", "unpack"))
```


### Load required packages 
```{r}
#| message: false
#| warning: false
library(kableExtra, exclude = "group_rows") # for displaying formatted tables w/ kbl()
library(janitor, include.only = c("clean_names", "tabyl"))
library(cowplot, include.only = "plot_grid")
library(tidyverse) # for general data wrangling
library(tidymodels) # for modeling
library(kknn)
library(discrim, exclude = "smoothness")
library(xfun, include.only = "cache_rds")
library(cowplot, include.only = "plot_grid")
library(readxl)
library(dplyr)
library(lubridate)
library(skimr)
library(tidyselect)
library(tibble)
library(effectsize)
```


### Source function scripts
```{r}
#| message: false
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_plots.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
source("C:/Users/16089/OneDrive/Desktop/lab psych610/610710_functions.R")
```


### Specify other global settings
```{r}
#| include: false
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max=Inf)
```

### file path
```{r}
file_path <- "C:/Users/1608/OneDrive/Desktop/Capstone_Project/semester_effect/raw_data_2"
```

```{r}
cleaned_path <- "C:/Users/16089/OneDrive/Desktop/Capstone_Project/semester_effect/cleaned_data_2"
```

## Cleaning EDA
### adjust the data columnnames 
#### for Questionnaire 
```{r}
data_1 <- read_csv(here::here(file_path, "Questionnaires_April 2, 2024_17.02_numbers.csv"), 
                   col_types = cols())
```

```{r}
data_2 <- read_csv(here::here(file_path, "Raven_April 2, 2024_16.11.csv"), 
                   col_types = cols())
```

```{r}
data_3 <- read_csv(here::here(file_path, "UCMRT_April 2, 2024_16.00.csv"), 
                   col_types = cols())
```

```{r}
headers <- colnames(data_1)
d_num <- read_csv(here::here(file_path, "Questionnaires_April 2, 2024_17.02_numbers.csv"), skip = 2, col_names = TRUE)
colnames(d_num) <- headers
```

```{r}
headers <- colnames(data_2)
d_num_2 <- read_csv(here::here(file_path, "Raven_April 2, 2024_16.11.csv"), skip = 2, col_names = TRUE)
colnames(d_num_2) <- headers
```

```{r}
headers <- colnames(data_3)
d_num_3 <- read_csv(here::here(file_path, "UCMRT_April 2, 2024_16.00.csv"), skip = 2, col_names = TRUE)
colnames(d_num_3) <- headers
```


### select the columns needed 
```{r}
d_num <- d_num %>%
  select(ParticipantID, EndDate, "SC0", "SC1", "SC2", "SC3", "SC4", "SC10", "SC11", "SC12", "SC13")
```

```{r}
d_num_2 <- d_num_2 %>% 
  select(ParticipantID, EndDate, Version, "SC0", Study, Q_TotalDuration)
```

```{r}
d_num_3 <- d_num_3 %>% 
  select(ParticipantID, EndDate, Version, "SC0", Study, Q_TotalDuration)
```


## Cleaning of the Raven
```{r}
d_num_2 <- d_num_2 %>%
  filter(str_detect(ParticipantID, "^1[0-9]{3}$")) %>% 
  filter(is.na(Study))

d_num <- d_num %>% 
  filter(str_detect(ParticipantID, "^1[0-9]{3}$"))

d_num_3 <- d_num_3 %>%
  filter(str_detect(ParticipantID, "^1[0-9]{3}$")) %>% 
  filter(is.na(Study))
```


```{r}
duplicate_rows <- d_num %>%
  group_by(ParticipantID) %>%
  filter(n() > 1) %>%
  ungroup()

# Print the duplicate rows
print(duplicate_rows)
```

```{r}
d_num <- d_num %>%
  group_by(ParticipantID) %>%                    # Group by participant_id
  filter(!if_any(starts_with("SC"), is.na) | n() == 1) %>%        # Remove rows with NA in response when duplicates exist
  ungroup()
```

only keep the first response from participant 1061
```{r}
d_num <- d_num %>%
  group_by(ParticipantID) %>%
  slice(1) %>%  # Keep only the first response for each participant
  ungroup()
```


### Renaming columns of three datasets
```{r}
d_num <- d_num %>%
  rename(participant_id_dataset = ParticipantID,
         date = EndDate)
```

```{r}
d_num_2 <- d_num_2 %>% 
  rename(participant_id_dataset = ParticipantID,
         date = EndDate,
         raven_hit_rate = SC0,
         raven_total_duration = Q_TotalDuration,
         raven_version = Version)
```

```{r}
d_num_3 <- d_num_3 %>% 
  rename(participant_id_dataset = ParticipantID,
         date = EndDate,
         ucmrt_hit_rate = SC0,
         ucmrt_total_duration = Q_TotalDuration,
         ucmrt_version = Version)
```

### Computing Scores for d_num
*For BFI score, divide by 2*
*Grit divide by 8*
*Conscientiousness divide by 10*
*Metacog divide by 30*
*Mindset divide by 20*

### BFI
```{r}
d_num$neuroticism_score <- d_num$SC3/2

d_num$conscientiousness_score <- d_num$SC2/2

d_num$openness_score <- d_num$SC4/2

d_num$extraversion_score <- d_num$SC0/2

d_num$agreeableness_score <- d_num$SC1/2
```

### Grit
```{r}
d_num$grit_score <- d_num$SC10/8
```

### Consciensciousness
```{r}
d_num$conscience_score <- d_num$SC11/10
```

### MetaCognition
```{r}
d_num$meta_cog <- d_num$SC12/30
```

### Mindset
```{r}
d_num$mindset_score <- d_num$SC13/20
```


### Computing Date
### adding time variables
#### year
```{r}
d_num$date <- trimws(d_num$date)
# Correct the format to match the original date format
d_num$date <- as.POSIXct(d_num$date, format = "%m/%d/%Y")

# Extract the year and create a new variable
d_num$year <- format(d_num$date, "%Y")

# Reorder columns
d_num <- d_num %>% select(participant_id_dataset, date, year, everything())

```


```{r}
unique(d_num$year)
```

#### semester_id

```{r}
# Define the function to categorize semesters
categorize_semester <- function(df) {
  
  # Print the first few dates to inspect the format
  print("Original Date Values:")
  print(head(df$date))
  
  # Check for unique date formats and values
  print("Unique Date Values:")
  print(unique(df$date))
  
  # Attempt to convert the date to Date format
  # This will only keep the date part (no time) and print rows that fail
  df$date <- as.Date(df$date, format = "%m/%d/%Y")
  
  # Identify rows where date conversion failed
  failed_rows <- df[is.na(df$date), ]
  if (nrow(failed_rows) > 0) {
    print("Rows where date conversion failed:")
    print(failed_rows)
    stop("Date conversion failed for some rows. Please check the format and values in the date column.")
  }
  
  # Create the semester_id variable based on the date
  df <- df %>%
    mutate(semester_id = case_when(
      date >= as.Date("2018-01-23") & date <= as.Date("2018-05-11") ~ "Spring-2018",
      date >= as.Date("2018-05-21") & date <= as.Date("2018-06-15") ~ "Summer-2018-4",
      date >= as.Date("2018-06-18") & date <= as.Date("2018-08-10") ~ "Summer-2018-8",
      date >= as.Date("2018-09-05") & date <= as.Date("2018-12-20") ~ "Fall-2018",
      date >= as.Date("2019-01-22") & date <= as.Date("2019-05-10") ~ "Spring-2019",
      date >= as.Date("2019-05-20") & date <= as.Date("2019-06-14") ~ "Summer-2019-4",
      date >= as.Date("2019-06-17") & date <= as.Date("2019-08-09") ~ "Summer-2019-8",
      date >= as.Date("2019-09-04") & date <= as.Date("2019-12-19") ~ "Fall-2019",
      date >= as.Date("2020-01-21") & date <= as.Date("2020-05-08") ~ "Spring-2020",
      date >= as.Date("2020-05-18") & date <= as.Date("2020-06-12") ~ "Summer-2020-4",
      date >= as.Date("2020-06-15") & date <= as.Date("2020-08-07") ~ "Summer-2020-8",
      date >= as.Date("2020-09-02") & date <= as.Date("2020-12-18") ~ "Fall-2020",
      date >= as.Date("2021-01-19") & date <= as.Date("2021-05-07") ~ "Spring-2021",
      date >= as.Date("2021-05-17") & date <= as.Date("2021-06-11") ~ "Summer-2021-4",
      date >= as.Date("2021-06-14") & date <= as.Date("2021-08-06") ~ "Summer-2021-8",
      date >= as.Date("2021-09-08") & date <= as.Date("2021-12-19") ~ "Fall-2021",
      date >= as.Date("2022-01-25") & date <= as.Date("2022-05-13") ~ "Spring-2022",
      date >= as.Date("2022-05-23") & date <= as.Date("2022-06-17") ~ "Summer-2022-4",
      date >= as.Date("2022-06-20") & date <= as.Date("2022-08-12") ~ "Summer-2022-8",
      date >= as.Date("2022-09-07") & date <= as.Date("2022-12-22") ~ "Fall-2022",
      date >= as.Date("2023-01-24") & date <= as.Date("2023-05-12") ~ "Spring-2023",
      date >= as.Date("2023-05-22") & date <= as.Date("2023-06-16") ~ "Summer-2023-4",
      date >= as.Date("2023-06-19") & date <= as.Date("2023-08-11") ~ "Summer-2023-8",
      date >= as.Date("2023-09-06") & date <= as.Date("2023-12-21") ~ "Fall-2023",
      TRUE ~ NA_character_
    ))

  return(df)
}

# Apply the function to your data frame d_num
d_num <- categorize_semester(d_num)

```


*Things to bring attention to, more data come from days between the official start and end dates of semesters; discuss on how to solve the gap, do we do it by changing the date? Or adding any new names?* 


```{r}
sum(is.na(d_num$semester_id))
```

#### academic_year 
```{r}
#(yyyy-xxxx)
#(yyyy-xxxx)
add_academic_year <- function(dataframe) {
  dataframe$academic_year <- NA 
  dataframe$academic_year[dataframe$semester_id %in% c("Spring-2018", "Summer-2018-4", "Summer-2018-8")] <- "2017-2018"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2018", "Spring-2019", "Summer-2019-4", "Summer-2019-8")] <- "2018-2019"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2019", "Spring-2020", "Summer-2020-4", "Summer-2020-8")] <- "2019-2020"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2020", "Spring-2021", "Summer-2021-4", "Summer-2021-8")] <- "2020-2021"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2021", "Spring-2022", "Summer-2022-4", "Summer-2022-8")] <- "2021-2022"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2022", "Spring-2023", "Summer-2023-4", "Summer-2023-8")] <- "2022-2023"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2023")] <- "2023-2024"
  return(dataframe)
}
d_num <- add_academic_year(d_num)
```

```{r}
sum(is.na(d_num$academic_year))
```

#### semester
```{r}
#(“Fall”, “Spring”, “Summer-4” and “Summer-8”)
add_semester <- function(dataframe) {
  # Initialize the new variable with NA
  dataframe$semester <- NA
  
  # Define the pattern and replacement rules
  dataframe$semester[grep("Fall", dataframe$semester_id)] <- "Fall"
  dataframe$semester[grep("Spring", dataframe$semester_id)] <- "Spring"
  dataframe$semester[grep("Summer", dataframe$semester_id)] <- "Summer"
  
  return(dataframe)
}

d_num <- add_semester(d_num)
```

#### first_day_of_the_semester
```{r}
add_first_day_of_the_semester <- function(dataframe) {
  # Define the mapping of semester_id to first day of the semester
  first_day_mapping <- list(
    "Spring-2018" = "2018-01-23",
    "Summer-2018-4" = "2018-05-21",
    "Summer-2018-8" = "2018-06-18",
    "Fall-2018" = "2018-09-05",
    "Spring-2019" = "2019-01-22",
    "Summer-2019-4" = "2019-05-20",
    "Summer-2019-8" = "2019-06-17",
    "Fall-2019" = "2019-09-04", 
    "Spring-2020" = "2020-01-21", 
    "Summer-2020-4" = "2020-05-18", 
    "Summer-2020-8" = "2020-06-15", 
    "Fall-2020" = "2020-09-02",
    "Spring-2021" = "2021-01-19", 
    "Summer-2021-4" = "2021-05-17", 
    "Summer-2021-8" = "2021-06-14", 
    "Fall-2021" = "2021-09-08", 
    "Spring-2022" = "2022-01-25", 
    "Summer-2022-4" = "2022-05-23", 
    "Summer-2022-8" = "2022-06-20", 
    "Fall-2022" = "2022-09-07", 
    "Spring-2023" = "2023-01-24", 
    "Summer-2023-4" = "2023-05-22", 
    "Summer-2023-8" = "2023-06-19", 
    "Fall-2023" = "2023-09-06"
  )
  
  # Assign first day of the semester based on semester_id
  dataframe$first_day_of_the_semester <- sapply(dataframe$semester_id, function(x) first_day_mapping[[x]])

  # Convert first_day_of_the_semester to character (in case it's not)
  dataframe$first_day_of_the_semester <- as.character(dataframe$first_day_of_the_semester)
  
  # Convert first_day_of_the_semester to Date format
  dataframe$first_day_of_the_semester <- as.Date(dataframe$first_day_of_the_semester, format = "%Y-%m-%d")
  
  # Check for the structure of the first_day_of_the_semester after conversion
  print(str(dataframe$first_day_of_the_semester))
  
  return(dataframe)
}

d_num <- add_first_day_of_the_semester(d_num)
```


#### days_since_semester_start
```{r}
add_days_since_semester_start <- function(dataframe) {
  # Convert the date column to Date format
  dataframe$date <- as.Date(dataframe$date, format = "%Y-%m-%d")
  
  # Convert first_day_of_the_semester to Date format (in case it's not already)
  dataframe$first_day_of_the_semester <- as.Date(dataframe$first_day_of_the_semester, format = "%Y-%m-%d")
  
  # Calculate days since semester start
  dataframe <- dataframe %>%
    mutate(days_since_semester_start = as.numeric(difftime(date, first_day_of_the_semester, units = "days")))
  
  return(dataframe)
}

d_num <- add_days_since_semester_start(d_num)
```

```{r}
d_num <- d_num |> 
  mutate(dataset_of_origin = "2021_JP_R56")
head(d_num)
```

```{r}
d_num <- d_num |> select(participant_id_dataset, semester_id, date, year, academic_year, semester,days_since_semester_start, first_day_of_the_semester, dataset_of_origin, neuroticism_score, conscientiousness_score, openness_score, extraversion_score, agreeableness_score, grit_score, conscience_score, meta_cog, mindset_score, everything())
```

-----Finished Cleaning d_num Questionnaire Data

## Adding time variables to the d_num_2 
### Computing Date

#### year
```{r}
# Correct the format to match the original date format
d_num_2$date <- as.Date(d_num_2$date, format = "%Y-%m-%d")

# Extract the year and create a new variable
d_num_2$year <- format(d_num_2$date, "%Y")

# Reorder columns
d_num_2 <- d_num_2 %>% select(participant_id_dataset, date, year, everything())
```


```{r}
unique(d_num_2$year)
```

#### semester_id

```{r}
# Define the function to categorize semesters
categorize_semester <- function(df) {
  
  # Print the first few dates to inspect the format
  print("Original Date Values:")
  print(head(df$date))
  
  # Check for unique date formats and values
  print("Unique Date Values:")
  print(unique(df$date))
  
  # Attempt to convert the date to Date format
  # This will only keep the date part (no time) and print rows that fail
  df$date <- as.Date(df$date, format = "%Y-%m-%d")
  
  # Identify rows where date conversion failed
  failed_rows <- df[is.na(df$date), ]
  if (nrow(failed_rows) > 0) {
    print("Rows where date conversion failed:")
    print(failed_rows)
    stop("Date conversion failed for some rows. Please check the format and values in the date column.")
  }
  
  # Create the semester_id variable based on the date
  df <- df %>%
    mutate(semester_id = case_when(
      date >= as.Date("2018-01-23") & date <= as.Date("2018-05-11") ~ "Spring-2018",
      date >= as.Date("2018-05-21") & date <= as.Date("2018-06-15") ~ "Summer-2018-4",
      date >= as.Date("2018-06-18") & date <= as.Date("2018-08-10") ~ "Summer-2018-8",
      date >= as.Date("2018-09-05") & date <= as.Date("2018-12-20") ~ "Fall-2018",
      date >= as.Date("2019-01-22") & date <= as.Date("2019-05-10") ~ "Spring-2019",
      date >= as.Date("2019-05-20") & date <= as.Date("2019-06-14") ~ "Summer-2019-4",
      date >= as.Date("2019-06-17") & date <= as.Date("2019-08-09") ~ "Summer-2019-8",
      date >= as.Date("2019-09-04") & date <= as.Date("2019-12-19") ~ "Fall-2019",
      date >= as.Date("2020-01-21") & date <= as.Date("2020-05-08") ~ "Spring-2020",
      date >= as.Date("2020-05-18") & date <= as.Date("2020-06-12") ~ "Summer-2020-4",
      date >= as.Date("2020-06-15") & date <= as.Date("2020-08-07") ~ "Summer-2020-8",
      date >= as.Date("2020-09-02") & date <= as.Date("2020-12-18") ~ "Fall-2020",
      date >= as.Date("2021-01-19") & date <= as.Date("2021-05-07") ~ "Spring-2021",
      date >= as.Date("2021-05-17") & date <= as.Date("2021-06-11") ~ "Summer-2021-4",
      date >= as.Date("2021-06-14") & date <= as.Date("2021-08-06") ~ "Summer-2021-8",
      date >= as.Date("2021-09-08") & date <= as.Date("2021-12-19") ~ "Fall-2021",
      date >= as.Date("2022-01-25") & date <= as.Date("2022-05-13") ~ "Spring-2022",
      date >= as.Date("2022-05-23") & date <= as.Date("2022-06-17") ~ "Summer-2022-4",
      date >= as.Date("2022-06-20") & date <= as.Date("2022-08-12") ~ "Summer-2022-8",
      date >= as.Date("2022-09-07") & date <= as.Date("2022-12-22") ~ "Fall-2022",
      date >= as.Date("2023-01-24") & date <= as.Date("2023-05-12") ~ "Spring-2023",
      date >= as.Date("2023-05-22") & date <= as.Date("2023-06-16") ~ "Summer-2023-4",
      date >= as.Date("2023-06-19") & date <= as.Date("2023-08-11") ~ "Summer-2023-8",
      date >= as.Date("2023-09-06") & date <= as.Date("2023-12-21") ~ "Fall-2023",
      TRUE ~ NA_character_
    ))

  return(df)
}

# Apply the function to your data frame d_num_2
d_num_2 <- categorize_semester(d_num_2)
```


*Things to bring attention to, more data come from days between the official start and end dates of semesters; discuss on how to solve the gap, do we do it by changing the date? Or adding any new names?* 


```{r}
sum(is.na(d_num_2$semester_id))
```

#### academic_year 
```{r}
#(yyyy-xxxx)
#(yyyy-xxxx)
add_academic_year <- function(dataframe) {
  dataframe$academic_year <- NA 
  dataframe$academic_year[dataframe$semester_id %in% c("Spring-2018", "Summer-2018-4", "Summer-2018-8")] <- "2017-2018"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2018", "Spring-2019", "Summer-2019-4", "Summer-2019-8")] <- "2018-2019"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2019", "Spring-2020", "Summer-2020-4", "Summer-2020-8")] <- "2019-2020"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2020", "Spring-2021", "Summer-2021-4", "Summer-2021-8")] <- "2020-2021"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2021", "Spring-2022", "Summer-2022-4", "Summer-2022-8")] <- "2021-2022"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2022", "Spring-2023", "Summer-2023-4", "Summer-2023-8")] <- "2022-2023"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2023")] <- "2023-2024"
  return(dataframe)
}
d_num_2 <- add_academic_year(d_num_2)
```

```{r}
sum(is.na(d_num_2$academic_year))
```

#### semester
```{r}
#(“Fall”, “Spring”, “Summer-4” and “Summer-8”)
add_semester <- function(dataframe) {
  # Initialize the new variable with NA
  dataframe$semester <- NA
  
  # Define the pattern and replacement rules
  dataframe$semester[grep("Fall", dataframe$semester_id)] <- "Fall"
  dataframe$semester[grep("Spring", dataframe$semester_id)] <- "Spring"
  dataframe$semester[grep("Summer", dataframe$semester_id)] <- "Summer"
  
  return(dataframe)
}

d_num_2 <- add_semester(d_num_2)
```

#### first_day_of_the_semester
```{r}
add_first_day_of_the_semester <- function(dataframe) {
  # Define the mapping of semester_id to first day of the semester
  first_day_mapping <- list(
    "Spring-2018" = "2018-01-23",
    "Summer-2018-4" = "2018-05-21",
    "Summer-2018-8" = "2018-06-18",
    "Fall-2018" = "2018-09-05",
    "Spring-2019" = "2019-01-22",
    "Summer-2019-4" = "2019-05-20",
    "Summer-2019-8" = "2019-06-17",
    "Fall-2019" = "2019-09-04", 
    "Spring-2020" = "2020-01-21", 
    "Summer-2020-4" = "2020-05-18", 
    "Summer-2020-8" = "2020-06-15", 
    "Fall-2020" = "2020-09-02",
    "Spring-2021" = "2021-01-19", 
    "Summer-2021-4" = "2021-05-17", 
    "Summer-2021-8" = "2021-06-14", 
    "Fall-2021" = "2021-09-08", 
    "Spring-2022" = "2022-01-25", 
    "Summer-2022-4" = "2022-05-23", 
    "Summer-2022-8" = "2022-06-20", 
    "Fall-2022" = "2022-09-07", 
    "Spring-2023" = "2023-01-24", 
    "Summer-2023-4" = "2023-05-22", 
    "Summer-2023-8" = "2023-06-19", 
    "Fall-2023" = "2023-09-06"
  )
  
  # Assign first day of the semester based on semester_id
  dataframe$first_day_of_the_semester <- sapply(dataframe$semester_id, function(x) first_day_mapping[[x]])

  # Convert first_day_of_the_semester to character (in case it's not)
  dataframe$first_day_of_the_semester <- as.character(dataframe$first_day_of_the_semester)
  
  # Convert first_day_of_the_semester to Date format
  dataframe$first_day_of_the_semester <- as.Date(dataframe$first_day_of_the_semester, format = "%Y-%m-%d")
  
  # Check for the structure of the first_day_of_the_semester after conversion
  print(str(dataframe$first_day_of_the_semester))
  
  return(dataframe)
}

d_num_2 <- add_first_day_of_the_semester(d_num_2)
```


#### days_since_semester_start
```{r}
add_days_since_semester_start <- function(dataframe) {
  # Convert the date column to Date format
  dataframe$date <- as.Date(dataframe$date, format = "%Y-%m-%d")
  
  # Convert first_day_of_the_semester to Date format (in case it's not already)
  dataframe$first_day_of_the_semester <- as.Date(dataframe$first_day_of_the_semester, format = "%Y-%m-%d")
  
  # Calculate days since semester start
  dataframe <- dataframe %>%
    mutate(days_since_semester_start = as.numeric(difftime(date, first_day_of_the_semester, units = "days")))
  
  return(dataframe)
}

d_num_2 <- add_days_since_semester_start(d_num_2)
```

```{r}
d_num_2 <- d_num_2 |> select(participant_id_dataset, semester_id, date, year, academic_year, semester,days_since_semester_start, first_day_of_the_semester, everything())
```

```{r}
d_num_2 <- d_num_2[, !names(d_num_2) %in% c("Study")]
```

----Finished Cleaning d_num_2

```{r}
d_num_2 %>%
  filter(participant_id_dataset == 1004) %>%
  print()
```

```{r}
d_num %>% 
  filter(participant_id_dataset == 1004) %>% 
  print()
```

```{r}
d_num_3 %>% 
  filter(participant_id_dataset == 1004) %>% 
  print()
```


## Cleaning of d_num_3 the UCMRT data
#### year
```{r}
# Correct the format to match the original date format
d_num_3$date <- as.Date(d_num_3$date, format = "%Y-%m-%d")

# Extract the year and create a new variable
d_num_3$year <- format(d_num_3$date, "%Y")

# Reorder columns
d_num_3 <- d_num_3 %>% select(participant_id_dataset, date, year, everything())
```


```{r}
unique(d_num_3$year)
```

#### semester_id

```{r}
# Define the function to categorize semesters
categorize_semester <- function(df) {
  
  # Print the first few dates to inspect the format
  print("Original Date Values:")
  print(head(df$date))
  
  # Check for unique date formats and values
  print("Unique Date Values:")
  print(unique(df$date))
  
  # Attempt to convert the date to Date format
  # This will only keep the date part (no time) and print rows that fail
  df$date <- as.Date(df$date, format = "%Y-%m-%d")
  
  # Identify rows where date conversion failed
  failed_rows <- df[is.na(df$date), ]
  if (nrow(failed_rows) > 0) {
    print("Rows where date conversion failed:")
    print(failed_rows)
    stop("Date conversion failed for some rows. Please check the format and values in the date column.")
  }
  
  # Create the semester_id variable based on the date
  df <- df %>%
    mutate(semester_id = case_when(
      date >= as.Date("2018-01-23") & date <= as.Date("2018-05-11") ~ "Spring-2018",
      date >= as.Date("2018-05-21") & date <= as.Date("2018-06-15") ~ "Summer-2018-4",
      date >= as.Date("2018-06-18") & date <= as.Date("2018-08-10") ~ "Summer-2018-8",
      date >= as.Date("2018-09-05") & date <= as.Date("2018-12-20") ~ "Fall-2018",
      date >= as.Date("2019-01-22") & date <= as.Date("2019-05-10") ~ "Spring-2019",
      date >= as.Date("2019-05-20") & date <= as.Date("2019-06-14") ~ "Summer-2019-4",
      date >= as.Date("2019-06-17") & date <= as.Date("2019-08-09") ~ "Summer-2019-8",
      date >= as.Date("2019-09-04") & date <= as.Date("2019-12-19") ~ "Fall-2019",
      date >= as.Date("2020-01-21") & date <= as.Date("2020-05-08") ~ "Spring-2020",
      date >= as.Date("2020-05-18") & date <= as.Date("2020-06-12") ~ "Summer-2020-4",
      date >= as.Date("2020-06-15") & date <= as.Date("2020-08-07") ~ "Summer-2020-8",
      date >= as.Date("2020-09-02") & date <= as.Date("2020-12-18") ~ "Fall-2020",
      date >= as.Date("2021-01-19") & date <= as.Date("2021-05-07") ~ "Spring-2021",
      date >= as.Date("2021-05-17") & date <= as.Date("2021-06-11") ~ "Summer-2021-4",
      date >= as.Date("2021-06-14") & date <= as.Date("2021-08-06") ~ "Summer-2021-8",
      date >= as.Date("2021-09-08") & date <= as.Date("2021-12-19") ~ "Fall-2021",
      date >= as.Date("2022-01-25") & date <= as.Date("2022-05-13") ~ "Spring-2022",
      date >= as.Date("2022-05-23") & date <= as.Date("2022-06-17") ~ "Summer-2022-4",
      date >= as.Date("2022-06-20") & date <= as.Date("2022-08-12") ~ "Summer-2022-8",
      date >= as.Date("2022-09-07") & date <= as.Date("2022-12-22") ~ "Fall-2022",
      date >= as.Date("2023-01-24") & date <= as.Date("2023-05-12") ~ "Spring-2023",
      date >= as.Date("2023-05-22") & date <= as.Date("2023-06-16") ~ "Summer-2023-4",
      date >= as.Date("2023-06-19") & date <= as.Date("2023-08-11") ~ "Summer-2023-8",
      date >= as.Date("2023-09-06") & date <= as.Date("2023-12-21") ~ "Fall-2023",
      TRUE ~ NA_character_
    ))

  return(df)
}

# Apply the function to your data frame d_num_3
d_num_3 <- categorize_semester(d_num_3)
```


*Things to bring attention to, more data come from days between the official start and end dates of semesters; discuss on how to solve the gap, do we do it by changing the date? Or adding any new names?* 


```{r}
sum(is.na(d_num_3$semester_id))
```

#### academic_year 
```{r}
#(yyyy-xxxx)
#(yyyy-xxxx)
add_academic_year <- function(dataframe) {
  dataframe$academic_year <- NA 
  dataframe$academic_year[dataframe$semester_id %in% c("Spring-2018", "Summer-2018-4", "Summer-2018-8")] <- "2017-2018"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2018", "Spring-2019", "Summer-2019-4", "Summer-2019-8")] <- "2018-2019"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2019", "Spring-2020", "Summer-2020-4", "Summer-2020-8")] <- "2019-2020"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2020", "Spring-2021", "Summer-2021-4", "Summer-2021-8")] <- "2020-2021"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2021", "Spring-2022", "Summer-2022-4", "Summer-2022-8")] <- "2021-2022"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2022", "Spring-2023", "Summer-2023-4", "Summer-2023-8")] <- "2022-2023"
  dataframe$academic_year[dataframe$semester_id %in% c("Fall-2023")] <- "2023-2024"
  return(dataframe)
}
d_num_3 <- add_academic_year(d_num_3)
```

```{r}
sum(is.na(d_num_3$academic_year))
```

#### semester
```{r}
#(“Fall”, “Spring”, “Summer-4” and “Summer-8”)
add_semester <- function(dataframe) {
  # Initialize the new variable with NA
  dataframe$semester <- NA
  
  # Define the pattern and replacement rules
  dataframe$semester[grep("Fall", dataframe$semester_id)] <- "Fall"
  dataframe$semester[grep("Spring", dataframe$semester_id)] <- "Spring"
  dataframe$semester[grep("Summer", dataframe$semester_id)] <- "Summer"
  
  return(dataframe)
}

d_num_3 <- add_semester(d_num_3)
```

#### first_day_of_the_semester
```{r}
add_first_day_of_the_semester <- function(dataframe) {
  # Define the mapping of semester_id to first day of the semester
  first_day_mapping <- list(
    "Spring-2018" = "2018-01-23",
    "Summer-2018-4" = "2018-05-21",
    "Summer-2018-8" = "2018-06-18",
    "Fall-2018" = "2018-09-05",
    "Spring-2019" = "2019-01-22",
    "Summer-2019-4" = "2019-05-20",
    "Summer-2019-8" = "2019-06-17",
    "Fall-2019" = "2019-09-04", 
    "Spring-2020" = "2020-01-21", 
    "Summer-2020-4" = "2020-05-18", 
    "Summer-2020-8" = "2020-06-15", 
    "Fall-2020" = "2020-09-02",
    "Spring-2021" = "2021-01-19", 
    "Summer-2021-4" = "2021-05-17", 
    "Summer-2021-8" = "2021-06-14", 
    "Fall-2021" = "2021-09-08", 
    "Spring-2022" = "2022-01-25", 
    "Summer-2022-4" = "2022-05-23", 
    "Summer-2022-8" = "2022-06-20", 
    "Fall-2022" = "2022-09-07", 
    "Spring-2023" = "2023-01-24", 
    "Summer-2023-4" = "2023-05-22", 
    "Summer-2023-8" = "2023-06-19", 
    "Fall-2023" = "2023-09-06"
  )
  
  # Assign first day of the semester based on semester_id
  dataframe$first_day_of_the_semester <- sapply(dataframe$semester_id, function(x) first_day_mapping[[x]])

  # Convert first_day_of_the_semester to character (in case it's not)
  dataframe$first_day_of_the_semester <- as.character(dataframe$first_day_of_the_semester)
  
  # Convert first_day_of_the_semester to Date format
  dataframe$first_day_of_the_semester <- as.Date(dataframe$first_day_of_the_semester, format = "%Y-%m-%d")
  
  # Check for the structure of the first_day_of_the_semester after conversion
  print(str(dataframe$first_day_of_the_semester))
  
  return(dataframe)
}

d_num_3 <- add_first_day_of_the_semester(d_num_3)
```


#### days_since_semester_start
```{r}
add_days_since_semester_start <- function(dataframe) {
  # Convert the date column to Date format
  dataframe$date <- as.Date(dataframe$date, format = "%Y-%m-%d")
  
  # Convert first_day_of_the_semester to Date format (in case it's not already)
  dataframe$first_day_of_the_semester <- as.Date(dataframe$first_day_of_the_semester, format = "%Y-%m-%d")
  
  # Calculate days since semester start
  dataframe <- dataframe %>%
    mutate(days_since_semester_start = as.numeric(difftime(date, first_day_of_the_semester, units = "days")))
  
  return(dataframe)
}

d_num_3 <- add_days_since_semester_start(d_num_3)
```

```{r}
d_num_3 <- d_num_3 |> select(participant_id_dataset, semester_id, date, year, academic_year, semester,days_since_semester_start, first_day_of_the_semester, everything())
```

```{r}
d_num_3 <- d_num_3[, !names(d_num_3) %in% c("Study")]
```

-----Finished Cleaning UCMRT tests data

## Merging all three datasets 
```{r}
d_num <- d_num %>% mutate(dataset = "questionnaires")
d_num_2 <- d_num_2 %>% mutate(dataset = "raven")
d_num_3 <- d_num_3 %>% mutate(dataset = "ucmrt")

# Combine datasets by rows with identifier
combined_data <- bind_rows(d_num, d_num_2, d_num_3)

```


## Final Output to CSV
```{r}
combined_data |> write_csv(here::here(cleaned_path, "R56_Combined_2.csv"))
```


# checking the LocationLongitude between 80- 100

*clean composite above, compute the date as before for the EDA of this data*

# Complie the R-56 datasets before Monday
# Upload to the folder
# Visualize the R-56 datasets
# Upload the images to the folder
# Write Down the Machine Learning Techniques I want to use
-why this technique(what is interesting about it)
-what info it can convey
# Write down the Machine Learning Notes 


