---
title: "Visualization_1"
author: "Shuheng Jiang"
date: "2024-10-18"
output: html_document
---

## Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Handle conflicts
```{r}
#| message: false
options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_ml.R?raw=true")
tidymodels_conflictRules()
# We also will need to resolve a new conflict using the following code
# Alternatively, John demonstrates code you can use when you load this library to prevent conflicts
conflictRules("Matrix", mask.ok = c("expand", "pack", "unpack"))
```

### Load required packages 
```{r}
#| message: false
#| warning: false
library(kableExtra, exclude = "group_rows") # for displaying formatted tables w/ kbl()
library(janitor, include.only = c("clean_names", "tabyl"))
library(cowplot, include.only = "plot_grid")
library(tidyverse) # for general data wrangling
library(tidymodels) # for modeling
library(kknn)
library(discrim, exclude = "smoothness")
library(xfun, include.only = "cache_rds")
library(cowplot, include.only = "plot_grid")
library(readxl)
library(dplyr)
library(lubridate)
library(skimr)
library(tidyselect)
library(tibble)
library(effectsize)
library(ggpubr)
devtools::install_github("NegativePotato/latltools")
library(latltools)
```

### Source function scripts
```{r}
#| message: false
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_plots.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
source("C:/Users/16089/OneDrive/Desktop/lab psych610/610710_functions.R")
```

### Specify other global settings
```{r}
#| include: false
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max=Inf)
```

### file path
```{r}
file_path <- "C:/Users/16089/OneDrive/Desktop/Capstone_Project/semester_effect/cleaned_data_postsurvey"
figure_path <- "C:/Users/16089/OneDrive/Desktop/Capstone_Project/semester_effect/figures_1"
```

### reading in data
```{r}
d_try <- read_csv(here::here(file_path, "Generic_Post_Survey_Combined.csv"), 
                   col_types = cols())
```


```{r}
range(d_try$days_since_semester_start)
```

### Semester effect on adhd_total

```{r}
mod_1 <- lm(adhd_total ~ days_since_semester_start, data = d_try)
anova(mod_1) %>% eta_squared()

plot_a_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = adhd_total)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Total score of 6-items ASRS', 
       title = 'Semester Effect on ADHD Total Score') +
  xlim(0, 110) +
  ylim(0, 7.0) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_a_complete, 
       filename = file.path(figure_path, 'plot_a_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_a_complete , file_name_full_no_ext = file.path(figure_path, "plot_a_complete"), width = 7, height = 7, units = 'in')
```


### Semester effect on ADHD_clinical
```{r}
glm_1 <- glm(adhd_clinical ~ days_since_semester_start, family = binomial(link = "logit"), data = d_try)
```

```{r}
newdata <- with(d_try, data.frame(
  days_since_semester_start = rep(unique(days_since_semester_start))
))
newdata$prob <- predict(glm_1, newdata = newdata, type = "response")
newdata$se <- predict(glm_1, newdata = newdata, type = "link", se.fit = TRUE)$se.fit

# Plot the figure
adhd_prob <- ggplot(d_try, aes(x = days_since_semester_start, y = adhd_clinical)) +
  geom_point(position = position_jitter(width = 0.1, height = 0.1), alpha = 0.5) +
  stat_smooth(data = newdata, aes(x = days_since_semester_start, y = prob, ymin = prob - 1.96 * se, ymax = prob + 1.96 * se),
              method = "glm", method.args = list(family = "binomial"),
              se = TRUE, fullrange = TRUE, alpha = 0.2) +
  scale_y_continuous(name = "Probability of being diagnosed ADHD", labels = scales::percent) +
  labs(title = "Semester Effect on Being Diagonized ADHD") +
  theme_bw() +
  stat_cor(color = 'blue', size = 5, label.x = 10, label.y = 0.8)
ggsave(plot = adhd_prob, 
       filename = file.path(figure_path, 'plot_adhd_prob.png'), 
       height = 6.5, width = 9, units = 'in')
```

```{r}
save_plot(plot = adhd_prob , file_name_full_no_ext = file.path(figure_path, "adhd_prob"), width = 7, height = 7, units = 'in')
```


### Semester effect on Agreeableness score
```{r}
mod_2 <- lm(agreeableness_score ~ days_since_semester_start, data = d_try)
anova(mod_2) %>% eta_squared()

plot_big5_agr_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = agreeableness_score)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Big-5 Agreeableness',
       title = "Semester Effect on Big-5 Agreeableness Score") +
  xlim(0, 110) +
  ylim(1, 5.5) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_big5_agr_complete, 
       filename = file.path(figure_path, 'plot_agr_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_big5_agr_complete, file_name_full_no_ext = file.path(figure_path, "plot_big5_agr_complete"), width = 7, height = 7, units = 'in')
```


### Semester effect on Conscientiousness score
```{r}
mod_3 <- lm(conscientiousness_score ~ days_since_semester_start, data = d_try)
anova(mod_3) %>% eta_squared()


plot_big5_con_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = conscientiousness_score)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Big-5 Conscientiousness',
       title = 'Semester Effect on Big-5 Conscientiousness Score') +
  xlim(0, 110) +
  ylim(1, 5.7) +
  stat_cor(color = "red", size = 6) 


ggsave(plot = plot_big5_con_complete, 
       filename = file.path(figure_path, 'plot_big5_con_complete.png'), 
       height = 5.5, width = 6,
       units = 'in')  
```

```{r}
save_plot(plot = plot_big5_con_complete, file_name_full_no_ext = file.path(figure_path, "plot_big5_con_complete"), width = 7, height = 7, units = 'in')
```


### Semester effect on Neuroticism score
```{r}
mod_4 <- lm(neuroticism_score ~ days_since_semester_start, data = d_try)
anova(mod_4) %>% eta_squared()

plot_big5_neu_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = neuroticism_score)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Big-5 Neuroticism',
       title = 'Semester Effect on Big-5 Neuroticism Score') +
  xlim(0,110)+
  ylim(1, 5.3) +
  stat_cor(color = "red", size = 6)

ggsave(plot = plot_big5_neu_complete, 
       filename = file.path(figure_path, 'plot_big5_neu_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_big5_neu_complete, file_name_full_no_ext = file.path(figure_path, "plot_big5_neu_complete"), width = 7, height = 7, units = 'in')
```

### Semester effect on Openness score
```{r}
mod_5 <- lm(openness_score ~ days_since_semester_start, data = d_try)
anova(mod_5) %>% eta_squared()


plot_big5_open_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = openness_score)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Big-5 Openness',
       title = 'Semester Effect on Big-5 Openness Score') +
  xlim(0,110)+
  ylim(1, 5.5) +
  stat_cor(color = "red", size = 6)

ggsave(plot = plot_big5_open_complete, 
       filename = file.path(figure_path, 'plot_big5_open_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_big5_open_complete, file_name_full_no_ext = file.path(figure_path, "plot_big5_open_complete"), width = 7, height = 7, units = 'in')
```


### Semester effect on Extraversion score
```{r}
mod_6 <- lm(extraversion_score ~ days_since_semester_start, data = d_try)
anova(mod_6) %>% eta_squared()


plot_big5_extraversion_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = extraversion_score)) +
  geom_jitter(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Big-5 Extraversion',
       title = 'Semester Effect on Big-5 Extraversion Score') +
  xlim(0,110)+
  ylim(1, 5.4) +
  stat_cor(color = 'red', size = 6)

ggsave(plot = plot_big5_extraversion_complete, 
       filename = file.path(figure_path, 'plot_big5_extraversion_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_big5_extraversion_complete , file_name_full_no_ext = file.path(figure_path, "plot_big5_extraversion_complete "), width = 7, height = 7, units = 'in')
```

### Semester effect on Growth Mindset Score
```{r}
mod_6 <- lm(growth_mindset_score ~ days_since_semester_start, data = d_try)
anova(mod_6) %>% eta_squared()

plot_growth_mindset_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = growth_mindset_score)) +
  geom_jitter(alpha = 0.6) +
  geom_smooth(method = "lm", color="red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Score for Growth Mindset',
       title = 'Semester Effect on Growth Mindset Score') +
  xlim(0,110) +
  ylim(1, 7.7) +
  stat_cor(color = 'red', size = 6)

ggsave(plot = plot_growth_mindset_complete, 
       filename = file.path(figure_path, 'plot_growth_mindset_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_growth_mindset_complete , file_name_full_no_ext = file.path(figure_path, "plot_growth_mindset_complete"), width = 7, height = 7, units = 'in')
```

### Semester effect on grit score
```{r}
mod_7 <- lm(grit_overall ~ days_since_semester_start, data = d_try)
anova(mod_7) %>% eta_squared()


plot_grit_overall_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = grit_overall)) +
  geom_jitter(alpah = 0.6) +
  geom_smooth(method = "lm", color="red", se = TRUE) +
  theme_bw(base_size = 14) +
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Score for Grit Overall',
       title = 'Semester Effect on Grit Score') +
  xlim(0,110) +
  ylim(1, 4.5) +
  stat_cor(color = 'red', size = 6)

ggsave(plot = plot_grit_overall_complete, 
       filename = file.path(figure_path, 'plot_grit_overall_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in') 
```

```{r}
save_plot(plot = plot_grit_overall_complete, file_name_full_no_ext = file.path(figure_path, "plot_grit_overall_complete"), width = 7, height = 7, units = 'in')
```


### Semester effect on grit interest 
```{r}
mod_8 <- lm(grit_interest ~ days_since_semester_start, data = d_try)
anova(mod_8) %>% eta_squared()


plot_grit_interest_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = grit_interest)) +
  geom_jitter() +
  geom_smooth(method = "lm", color="red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Score for Grit Interesr',
       title = 'Semester Effect on Grit Interest Score') +
  xlim(0,110)+
  ylim(1, 4.5) +
  stat_cor(color = "red", size = 6)

ggsave(plot = plot_grit_interest_complete, 
       filename = file.path(figure_path, 'plot_grit_interest_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```

```{r}
save_plot(plot = plot_grit_interest_complete, file_name_full_no_ext = file.path(figure_path, "plot_grit_interest_complete"), width = 7, height = 7, units = 'in')
```


### Semester effect on grit effort
```{r}
mod_9 <- lm(grit_effort ~ days_since_semester_start, data = d_try)
anova(mod_9) %>% eta_squared()


plot_grit_effort_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = grit_effort)) +
  geom_jitter(alpah = 0.6) +
  geom_smooth(method = "lm", color="red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Score for Grit Effort',
       title = 'Semester Effect on Grit Effort Score') +
  xlim(0,110)+
  ylim(1, 4.5) +
  stat_cor(color = "red", size = 6)

ggsave(plot = plot_grit_effort_complete, 
       filename = file.path(figure_path, 'plot_grit_effort_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_grit_effort_complete, file_name_full_no_ext = file.path(figure_path, "plot_grit_effort_complete"), width = 7, height = 7, units = 'in')
```


### Semester effect on Overall Matrices Hit Rate 
```{r}
mod_10 <- lm(matrices_hit_rate ~ days_since_semester_start, data = d_try)
anova(mod_10) %>% eta_squared()

plot_matrices_hit_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = matrices_hit_rate)) +
  geom_jitter() +
  geom_smooth(method = "lm", color="red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Matrices Hit Rate',
       title = 'Semester Effect on Matrices Hit Rate') +
  xlim(0,110)+
  ylim(0, 1.2) +
  stat_cor(color = "red", size = 6)

ggsave(plot = plot_matrices_hit_complete, 
       filename = file.path(figure_path, 'plot_matrices_hit_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_matrices_hit_complete, file_name_full_no_ext = file.path(figure_path, "plot_matrices_hit_complete"), width = 7, height = 7, units = 'in')
```

### Semester effect on Sandia Matrices Hit Rate 
```{r}
mod_11 <- lm(sandia_matrices_hit_rate ~ days_since_semester_start, data = d_try)
anova(mod_11) %>% eta_squared()


d_graph <- data.frame(days_since_semester_start = seq(0, 110, length = 1000))
d_graph <- ggplotPredict(mod_11, d_graph)


plot_sandiamatrices_hit_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = sandia_matrices_hit_rate)) +
  geom_jitter() +
  geom_smooth(data = d_graph, aes(ymin = CILo, ymax = CIHi, x = days_since_semester_start, y =Predicted),
              stat = "identity", color="red") +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', y = 'Sandia Matrices Hit Rate') +
  xlim(0,110)+
  ylim(0, 1) +
  stat_cor()
```

```{r}
save_plot(plot = plot_sandiamatrices_hit_complete, file_name_full_no_ext = file.path(figure_path, "plot_sandiamatrices_hit_complete"), width = 7, height = 7, units = 'in')
```

### Semester Effect in ucmrt hit rate
```{r}
mod_12 <- lm(ucmrt_hit_rate ~ days_since_semester_start, data = d_try)
anova(mod_12) %>% eta_squared()


d_graph <- data.frame(days_since_semester_start = seq(0, 110, length = 1000))
d_graph <- ggplotPredict(mod_12, d_graph)


plot_ucmrt_hit_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = ucmrt_hit_rate)) +
  geom_jitter() +
  geom_smooth(data = d_graph, aes(ymin = CILo, ymax = CIHi, x = days_since_semester_start, y =Predicted),
              stat = "identity", color="red") +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', y = 'UCMRT Hit Rate') +
  xlim(0,110)+
  ylim(0, 1) +
  stat_cor()
plot_ucmrt_hit_complete
```

```{r}
save_plot(plot = plot_ucmrt_hit_complete, file_name_full_no_ext = file.path(figure_path, "plot_ucmrt_hit_complete"), width = 7, height = 7, units = 'in')
```
