---
title: "Visualization_3"
author: "Shuheng Jiang"
date: "2024-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Handle conflicts
```{r}
#| message: false
options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_ml.R?raw=true")
tidymodels_conflictRules()
# We also will need to resolve a new conflict using the following code
# Alternatively, John demonstrates code you can use when you load this library to prevent conflicts
conflictRules("Matrix", mask.ok = c("expand", "pack", "unpack"))
```

### Load required packages 
```{r}
#| message: false
#| warning: false
library(kableExtra, exclude = "group_rows") # for displaying formatted tables w/ kbl()
library(janitor, include.only = c("clean_names", "tabyl"))
library(cowplot, include.only = "plot_grid")
library(tidyverse) # for general data wrangling
library(tidymodels) # for modeling
library(kknn)
library(discrim, exclude = "smoothness")
library(xfun, include.only = "cache_rds")
library(cowplot, include.only = "plot_grid")
library(readxl)
library(dplyr)
library(lubridate)
library(skimr)
library(tidyselect)
library(tibble)
library(effectsize)
library(ggpubr)
devtools::install_github("NegativePotato/latltools")
library(latltools)
```

### Source function scripts
```{r}
#| message: false
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_plots.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
source("C:/Users/16089/OneDrive/Desktop/lab psych610/610710_functions.R")
```

### Specify other global settings
```{r}
#| include: false
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max=Inf)
```

### file path
```{r}
file_path <- "C:/Users/16089/OneDrive/Desktop/Capstone_Project/semester_effect/cleaned_data_3"
figure_path <- "C:/Users/16089/OneDrive/Desktop/Capstone_Project/semester_effect/figures_3_personality"
figure_path_2 <- "C:/Users/16089/OneDrive/Desktop/Capstone_Project/semester_effect/figures_3_matrices"
figure_path_3 <- "C:/Users/16089/OneDrive/Desktop/Capstone_Project/semester_effect/figures_3_torr"
```

### reading in data
```{r}
d_personality <- read_csv(here::here(file_path, "Clean_2021_EY_FYP_Personality.csv"), 
                   col_types = cols())
```

```{r}
d_torr <- read_csv(here::here(file_path, "Clean_2022_LA_Patterns_FYP_torr.csv"), 
                   col_types = cols())
```

```{r}
d_matrices <- read_csv(here::here(file_path, "Clean_AC_2021_wmgf_matrices.csv"), 
                   col_types = cols())
```


## Visualization of Clean_2021_EY_FYP_Personality


### semester effect on extraversion score

```{r}
range(d_personality$days_since_semester_start)
```

```{r}
range(d_personality$extraversion)
```

```{r}
mod_1 <- lm(extraversion ~ days_since_semester_start, data = d_personality)
anova(mod_1) %>% eta_squared()

plot_extraversion_complete <- ggplot(d_personality, aes(x = days_since_semester_start, y = extraversion)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Extraversion Score', 
       title = 'Semester Effect on Extraversion Score') +
  xlim(60, 110) +
  ylim(15, 36) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_extraversion_complete, 
       filename = file.path(figure_path, 'plot_extraversion_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```

```{r}
save_plot(plot = plot_extraversion_complete, file_name_full_no_ext = file.path(figure_path, "plot_extraversion_complete"), width = 7, height = 7, units = 'in')
```


### semester effect on agreeableness score

```{r}
range(d_personality$agreeableness)
```

```{r}
mod_2 <- lm(agreeableness ~ days_since_semester_start, data = d_personality)
anova(mod_2) %>% eta_squared()

plot_agreeableness_complete <- ggplot(d_personality, aes(x = days_since_semester_start, y = agreeableness)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Agreeableness Score', 
       title = 'Semester Effect on Agreeableness Score') +
  xlim(60, 110) +
  ylim(17, 45) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_agreeableness_complete, 
       filename = file.path(figure_path, 'plot_agreeableness_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```

```{r}
save_plot(plot = plot_agreeableness_complete, file_name_full_no_ext = file.path(figure_path, "plot_agreeableness_complete"), width = 7, height = 7, units = 'in')
```

### semester effect on conscientiousness score
```{r}
range(d_personality$conscientiousness)
```

The p is significant and eta square is larger 
```{r}
mod_3 <- lm(conscientiousness ~ days_since_semester_start, data = d_personality)
anova(mod_3) %>% eta_squared()

plot_conscientiousness_complete <- ggplot(d_personality, aes(x = days_since_semester_start, y = conscientiousness)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Conscientiousness Score', 
       title = 'Semester Effect on Conscientiousness Score') +
  xlim(60, 110) +
  ylim(18, 45) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_conscientiousness_complete, 
       filename = file.path(figure_path, 'plot_conscientiousness_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```

```{r}
save_plot(plot = plot_conscientiousness_complete, file_name_full_no_ext = file.path(figure_path, "plot_conscientiousness_complete"), width = 7, height = 7, units = 'in')
```

### semester effect on neuroticism score
```{r}
range(d_personality$neuroticism)
```

```{r}
mod_4 <- lm(neuroticism ~ days_since_semester_start, data = d_personality)
anova(mod_4) %>% eta_squared()

plot_neuroticism_complete <- ggplot(d_personality, aes(x = days_since_semester_start, y = neuroticism)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Neuroticism Score', 
       title = 'Semester Effect on Neuroticism Score') +
  xlim(60, 110) +
  ylim(9, 40) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_neuroticism_complete, 
       filename = file.path(figure_path, 'plot_neuroticism_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```

```{r}
save_plot(plot = plot_neuroticism_complete, file_name_full_no_ext = file.path(figure_path, "plot_neuroticism_complete"), width = 7, height = 7, units = 'in')
```

### semester effect on openness
```{r}
range(d_personality$openness)
```

```{r}
mod_5 <- lm(openness ~ days_since_semester_start, data = d_personality)
anova(mod_5) %>% eta_squared()

plot_openness_complete <- ggplot(d_personality, aes(x = days_since_semester_start, y = openness)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Openness Score', 
       title = 'Semester Effect on Openness Score') +
  xlim(60, 110) +
  ylim(14, 50) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_openness_complete, 
       filename = file.path(figure_path, 'plot_openness_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```

```{r}
save_plot(plot = plot_openness_complete, file_name_full_no_ext = file.path(figure_path, "plot_openness_complete"), width = 7, height = 7, units = 'in')
```

### semester effect on metacognition
```{r}
mod_6 <- lm(metacognition ~ days_since_semester_start, data = d_personality)
anova(mod_6) %>% eta_squared()

plot_meta_complete <- ggplot(d_personality, aes(x = days_since_semester_start, y = metacognition)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'MetaCognition Score', 
       title = 'Semester Effect on MetaCognition Score') +
  xlim(60, 110) +
  ylim(30, 100) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_meta_complete, 
       filename = file.path(figure_path, 'plot_meta_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```


```{r}
save_plot(plot = plot_meta_complete, file_name_full_no_ext = file.path(figure_path, "plot_meta_complete"), width = 7, height = 7, units = 'in')
```

### semester effect on metacognitive_motivations

```{r}
range(d_personality$metacognitive_motivations)
```

```{r}
mod_7 <- lm(metacognitive_motivations ~ days_since_semester_start, data = d_personality)
anova(mod_7) %>% eta_squared()

plot_metamotive_complete <- ggplot(d_personality, aes(x = days_since_semester_start, y = metacognitive_motivations)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Metacognitive Motivations Score', 
       title = 'Semester Effect on MetaCognitive Motivations Score') +
  xlim(60, 110) +
  ylim(10, 50) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_metamotive_complete, 
       filename = file.path(figure_path, 'plot_metamotive_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_metamotive_complete, file_name_full_no_ext = file.path(figure_path, "plot_metamotive_complete"), width = 7, height = 7, units = 'in')
```


### semester effect on metacognitive_skills
```{r}
range(d_personality$metacognitive_skills)
```

```{r}
mod_8 <- lm(metacognitive_skills ~ days_since_semester_start, data = d_personality)
anova(mod_8) %>% eta_squared()

plot_metaskills_complete <- ggplot(d_personality, aes(x = days_since_semester_start, y = metacognitive_skills)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Metacognitive Skills Score', 
       title = 'Semester Effect on MetaCognitive Skills Score') +
  xlim(60, 110) +
  ylim(10, 60) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_metaskills_complete, 
       filename = file.path(figure_path, 'plot_metaskills_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```


```{r}
save_plot(plot = plot_metaskills_complete, file_name_full_no_ext = file.path(figure_path, "plot_metaskills_complete"), width = 7, height = 7, units = 'in')
```

## Visualization of Clean_AC_2021_wmgf_matrices

### semester effect on ucmrt score

```{r}
mod_9 <- lm(ucmrt_hit_rate ~ days_since_semester_start, data = d_matrices)
anova(mod_9) %>% eta_squared()

plot_ucmrt_complete <- ggplot(d_matrices, aes(x = days_since_semester_start, y = ucmrt_hit_rate)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'UCMRT Hit Score', 
       title = 'Semester Effect on UCMRT Score') +
  xlim(0, 110) +
  ylim(0, 1) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_ucmrt_complete, 
       filename = file.path(figure_path_2, 'plot_ucmrt_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_ucmrt_complete, file_name_full_no_ext = file.path(figure_path_2, "plot_ucmrt_complete"), width = 7, height = 7, units = 'in')
```

### semester effect on sandia score
```{r}
mod_10 <- lm(sandia_matrices_hit_rate ~ days_since_semester_start, data = d_matrices)
anova(mod_10) %>% eta_squared()

plot_sandia_complete <- ggplot(d_matrices, aes(x = days_since_semester_start, y = sandia_matrices_hit_rate)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Sandia Hit Score', 
       title = 'Semester Effect on Sandia Score') +
  xlim(0, 110) +
  ylim(0, 1) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_sandia_complete, 
       filename = file.path(figure_path_2, 'plot_sandia_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_sandia_complete, file_name_full_no_ext = file.path(figure_path_2, "plot_sandia_complete"), width = 7, height = 7, units = 'in')
```

### semester effect on overall matrices score

```{r}
mod_11 <- lm(matrices_hit_rate ~ days_since_semester_start, data = d_matrices)
anova(mod_11) %>% eta_squared()

plot_matrices_complete <- ggplot(d_matrices, aes(x = days_since_semester_start, y = matrices_hit_rate)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Matrices Hit Score', 
       title = 'Semester Effect on Matrices Score') +
  xlim(0, 110) +
  ylim(0, 1) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_matrices_complete, 
       filename = file.path(figure_path_2, 'plot_matrices_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_matrices_complete, file_name_full_no_ext = file.path(figure_path_2, "plot_matrices_complete"), width = 7, height = 7, units = 'in')
```

## Visualization of Clean_2022_LA_Patterns_FYP_torr

### semester effect on Realtional Reasoning Set 1 Score

```{r}
range(d_torr$torr_set1_score)
```
```{r}
range(d_torr$days_since_semester_start)
```


```{r}
mod_12 <- lm(torr_set1_score ~ days_since_semester_start, data = d_torr)
anova(mod_12) %>% eta_squared()

plot_torr1_complete <- ggplot(d_torr, aes(x = days_since_semester_start, y = torr_set1_score)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Relational Reasoning Set 1 Score', 
       title = 'Semester Effect on Relational Reasoning Score') +
  xlim(0, 100) +
  ylim(0, 8) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_torr1_complete, 
       filename = file.path(figure_path_3, 'plot_torr1_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_torr1_complete, file_name_full_no_ext = file.path(figure_path_3, "plot_torr1_complete"), width = 7, height = 7, units = 'in')
```

### semester effect on Realtional Reasoning Set 2 Score

```{r}
range(d_torr$torr_set2_score)
```

```{r}
mod_13 <- lm(torr_set2_score ~ days_since_semester_start, data = d_torr)
anova(mod_13) %>% eta_squared()

plot_torr2_complete <- ggplot(d_torr, aes(x = days_since_semester_start, y = torr_set2_score)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Relational Reasoning Set 2 Score', 
       title = 'Semester Effect on Relational Reasoning Score') +
  xlim(0, 100) +
  ylim(0, 8) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_torr2_complete, 
       filename = file.path(figure_path_3, 'plot_torr2_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```


```{r}
save_plot(plot = plot_torr2_complete, file_name_full_no_ext = file.path(figure_path_3, "plot_torr2_complete"), width = 7, height = 7, units = 'in')
```

### semester effect on atmi score

```{r}
range(d_torr$atmi_score)
```

```{r}
mod_14 <- lm(atmi_score ~ days_since_semester_start, data = d_torr)
anova(mod_14) %>% eta_squared()

plot_atmi_complete <- ggplot(d_torr, aes(x = days_since_semester_start, y = atmi_score)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Attitude towards Maths Score', 
       title = 'Semester Effect on Attitude towards Maths') +
  xlim(0, 100) +
  ylim(48, 132) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_atmi_complete, 
       filename = file.path(figure_path_3, 'plot_atmi_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
save_plot(plot = plot_atmi_complete, file_name_full_no_ext = file.path(figure_path_3, "plot_atmi_complete"), width = 7, height = 7, units = 'in')
```

### semester effect on wj score
```{r}
range(d_torr$wj_score)
```

```{r}
mod_15 <- lm(wj_score ~ days_since_semester_start, data = d_torr)
anova(mod_15) %>% eta_squared()

plot_wj_complete <- ggplot(d_torr, aes(x = days_since_semester_start, y = wj_score)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(x = 'Days since semester starts when completed the survey', 
       y = 'Maths Fluency Test Score', 
       title = 'Semester Effect on Math Fluency Test') +
  xlim(0, 100) +
  ylim(38, 160) +
  stat_cor(color = "red", size = 6) 

ggsave(plot = plot_wj_complete, 
       filename = file.path(figure_path_3, 'plot_wj_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```


```{r}
save_plot(plot = plot_wj_complete, file_name_full_no_ext = file.path(figure_path_3, "plot_wj_complete"), width = 7, height = 7, units = 'in')
```


