---
title: "Models_Testing"
author: "Shuheng Jiang"
date: "2024-10-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
### Handle conflicts
```{r}
#| message: false
options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_ml.R?raw=true")
tidymodels_conflictRules()
conflictRules("Matrix", mask.ok = c("expand", "pack", "unpack"))
```


### Load required packages 
```{r}
#| message: false
#| warning: false
library(kableExtra, exclude = "group_rows") # for displaying formatted tables w/ kbl()
library(janitor, include.only = c("clean_names", "tabyl"))
library(cowplot, include.only = "plot_grid")
library(tidyverse) # for general data wrangling
library(tidymodels) # for modeling
library(kknn)
library(discrim, exclude = "smoothness")
library(xfun, include.only = "cache_rds")
library(cowplot, include.only = "plot_grid")
library(readxl)
library(dplyr)
library(lubridate)
library(skimr)
library(tidyselect)
library(tibble)
library(effectsize)
library(lme4)
library(ggplot2)
library(effects)
library(nnet)
library(broom)
library(lavaan)
library(ggeffects)
```


### Source function scripts
```{r}
#| message: false
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_plots.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
source("C:/Users/16089/OneDrive/Desktop/lab psych610/610710_functions.R")
```

### Specify other global settings
```{r}
#| include: false
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max=Inf)
```

### file path
```{r}
file_path <- "C:/Users/16089/OneDrive/Desktop/Greens lab/semester/raw_data"
```

```{r}
d_try <- read_csv(here::here(file_path, "Generic_Post_Survey_Combined.csv"), 
                   col_types = cols())
```


## Mediation
```{r}
require(lavaan)
library(beepr)
```


```{r}
d_try |> 
  select(days_since_semester_start, grit_overall, grit_interest, grit_effort, matrices_hit_rate, sandia_matrices_hit_rate, ucmrt_hit_rate, neuroticism_score, extraversion_score, conscientiousness_score, openness_score, agreeableness_score, growth_mindset_score, adhd_total, adhd_clinical) |> 
  cor(use = "pairwise.complete.obs") |> 
  corrplot::corrplot.mixed(tl.pos = "lt")
```


### Mediation model 1 
Let's verbally describe our hypothesized mediation model.
>> Higher Growth Mindset scores leads to higher self-control (lower grit score), which in turn leads to students choose to take the survey earlier in the semester.

#### X = ADHD_clinical
#### M = conscientious
#### Y = days_since_start_of_the_semester


Explanation: conscientiousness construct overlap with self_control;

```{r}
range(d_try$days_since_semester_start)
```

```{r}
hist(d_try$days_since_semester_start)
```


Need to Apply Log transformation, it is highly negatively skewed 
```{r}
d_try$days_since_semester_start_log <- log(d_try$days_since_semester_start)
```

# 1. There is a relationship between X and Y. Path c is significant.
```{r}
m1 <- lm(days_since_semester_start_log ~ adhd_clinical, data = d_try)
summary(m1)
```


```{r}
m2 <- lm(conscientiousness_score ~ adhd_clinical, data = d_try)
summary(m2)
```


```{r}
m3 <- lm(days_since_semester_start_log ~ conscientiousness_score + adhd_clinical, data = d_try)
summary(m3)
```
Yes! There is a mediation effect to explore; But the hypothesis might not be valid, since adhd is a clinical diagnosis and conscientiousness_score is a personality test score. 


Test whether c' is significantly reduced when introducing a*b
```{r}
medmodel_1 <- ' 

# Lets first specify the models, and we can calculate other effects.
# We want to predict conscientiousness(self_control) from adhd diagnosis. AND
# We want to predict the day that come to test from self_control and adhd diagnosis.

conscientiousness_score ~ A * adhd_clinical
days_since_semester_start_log ~ CPRIME * adhd_clinical + B * conscientiousness_score

# By adding "A" and "CPRIME" we are labeling the coefficients.

# Define effects
AB:= A*B # Indirect effect
Total := CPRIME + A*B # Total effect
'
```


```{r}
set.seed(20010429)
fit2 <- sem(medmodel_1, data = d_try, se = "bootstrap", bootstrap = 1000)
parameterEstimates(fit2, ci = TRUE, level = 0.95, boot.ci.type = "perc", zstat = F)
```


```{r}
varTable(fit2)
```

The Confidence Interval Does Not include 0 for path AB; Condition 4 of the mediation test is satisfied. C' is now not significant.


### Medaition Model 2 
#### X = ADHD_clinical
#### M = grit 
#### Y = days_since_start_of_the_semester
```{r}
# New mediation model with grit score as the mediator (M)
medmodel_grit <- '
  grit_overall ~ A * adhd_clinical
  days_since_semester_start_log ~ CPRIME * adhd_clinical + B * grit_overall
  
  # Indirect effect
  AB := A * B
  # Total effect
  Total := CPRIME + A * B
'

set.seed(20010429)
fit_grit <- sem(medmodel_grit, data = d_try, se = "bootstrap", bootstrap = 1000)
parameterEstimates(fit_grit, ci = TRUE, level = 0.95, boot.ci.type = "perc", zstat = FALSE)
```
There does not exist a mediation effect 


### Medaition Model 3
#### X = 
#### M = 
#### Y = days_since_start_of_the_semester
```{r}
# Specify the mediation model with grit as the mediator
medmodel_grit <- '
  grit_overall ~ A * ucmrt_hit_rate  # Path A: Does fluid intelligence predict grit?
  days_since_semester_start_log ~ CPRIME * ucmrt_hit_rate + B * grit_overall 

  # Indirect effect (A * B) and Total effect (CPRIME + A * B)
  AB := A * B  # Indirect effect
  Total := CPRIME + A * B  # Total effect
'

# Fit the model
set.seed(20010429)
fit_grit <- sem(medmodel_grit, data = d_try, se = "bootstrap", bootstrap = 1000)

# Output parameter estimates
parameterEstimates(fit_grit, ci = TRUE, level = 0.95, boot.ci.type = "perc", zstat = FALSE)
```

No Mediation Effect 


