---
title: "Visualization_Presentation"
author: "Shuheng Jiang"
date: "2024-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Handle conflicts
```{r}
#| message: false
options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_ml.R?raw=true")
tidymodels_conflictRules()
# We also will need to resolve a new conflict using the following code
# Alternatively, John demonstrates code you can use when you load this library to prevent conflicts
conflictRules("Matrix", mask.ok = c("expand", "pack", "unpack"))
```

### Load required packages 
```{r}
#| message: false
#| warning: false
library(kableExtra, exclude = "group_rows") # for displaying formatted tables w/ kbl()
library(janitor, include.only = c("clean_names", "tabyl"))
library(cowplot, include.only = "plot_grid")
library(tidyverse) # for general data wrangling
library(tidymodels) # for modeling
library(kknn)
library(discrim, exclude = "smoothness")
library(xfun, include.only = "cache_rds")
library(cowplot, include.only = "plot_grid")
library(readxl)
library(dplyr)
library(lubridate)
library(skimr)
library(tidyselect)
library(tibble)
library(effectsize)
library(ggpubr)
devtools::install_github("NegativePotato/latltools")
library(latltools)
```

### Source function scripts
```{r}
#| message: false
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_plots.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
source("C:/Users/16089/OneDrive/Desktop/lab psych610/610710_functions.R")
```

### Specify other global settings
```{r}
#| include: false
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max=Inf)
```

### file path
```{r}
file_path <- "C:/Users/16089/OneDrive/Desktop/Capstone_Project/semester_effect/cleaned_data_1"
file_path_2 <- "C:/Users/16089/OneDrive/Desktop/Capstone_Project/semester_effect/cleaned_data_3"
figure_path <- "C:/Users/16089/OneDrive/Desktop/Capstone_Project/semester_effect/figures_to_presentation"
```



### reading in data


```{r}
d_try <- read_csv(here::here(file_path, "Generic_Post_Survey_Combined.csv"), 
                   col_types = cols())
```

```{r}
d_personality <- read_csv(here::here(file_path_2, "Clean_2021_EY_FYP_Personality.csv"), 
                   col_types = cols())
```

```{r}
d_matrices <- read_csv(here::here(file_path_2, "Clean_AC_2021_wmgf_matrices.csv"), 
                   col_types = cols())
```

## Histogram
```{r}
d_try$semester_id <- factor(d_try$semester_id, 
                            levels = c("Spring-2018", "Fall-2018", "Spring-2019", 
                                       "Fall-2019", "Spring-2020", "Fall-2020", 
                                       "Spring-2021", "Fall-2021", "Spring-2022", 
                                       "Fall-2022", "Fall-2023"))

ggplot(data = d_try) + 
  geom_bar(aes(x= days_since_semester_start)) +
  facet_wrap(~semester_id) 
  
```



```{r}
ggplot(data = d_try) + 
  geom_bar(aes(x = days_since_semester_start), fill = "red") + # Adjust bar colors
  facet_wrap(~semester_id) +  # Arrange facets in a grid
  theme_minimal() +  # Use a clean theme
  labs(
    title = "Enrollment Distribution of Participants By Semester",
    x = "Days Since Semester Start",
    y = "Enrollment Frequency",
    caption = "Data source: Data_PostSurvey"
  ) 
```

```{r}

ggplot(data = d_try) + 
  geom_bar(aes(x= days_since_semester_start), fill="red") +
  theme_minimal() +  # Use a clean theme
  labs(
    title = "Enrollment Distribution of Participants",
    x = "Days Since Semester Start",
    y = "Enrollment Frequency",
    caption = "Data source: Data_PostSurvey"
  ) 
```

```{r}
unique(d_matrices$semester_id)
```
```{r}
d_matrices$semester_id <- factor(d_matrices$semester_id, 
                            levels = c("fall_2018", "spring_2019", "summer-8_2019",
                                       "fall_2019", "spring_2020"))

ggplot(data = d_matrices) + 
  geom_bar(aes(x= days_since_semester_start), fill="brown")+
  theme_minimal() +  # Use a clean theme
  labs(
    title = "Enrollment Distribution of Participants",
    x = "Days Since Semester Start",
    y = "Enrollment Frequency",
    caption = "Data source: Data_Matrices"
  ) 
```

```{r}
distribution_4 <- ggplot(data = d_personality) + 
  geom_bar(aes(x= days_since_semester_start), fill="blue")+
    theme_minimal() +  # Use a clean theme
    labs(
    title = "Enrollment Distribution of Participants",
    x = "Days Since Semester Start",
    y = "Enrollment Frequency",
    caption = "Data source: Data_Personality"
  ) 
distribution_4
```



## Interaction
```{r}
ggplot() +
  geom_point(data = d_try, aes(x = matrices_hit_rate, y = days_since_semester_start, color = semester), alpha = 0.5) +
  geom_smooth(data = d_try, aes(x = matrices_hit_rate, y = days_since_semester_start, color = semester), method = "lm")
```


```{r}
library(dplyr)

```

```{r}
ggplot() +
  geom_point(data = d_try, aes(x = days_since_semester_start, y = matrices_hit_rate, color = semester), alpha = 0.5) +
  geom_smooth(data = d_try, aes(x = days_since_semester_start, y = matrices_hit_rate, color = semester), method = "lm")
```


```{r}
# Convert 0 and 1 to "no" and "yes"
d_try$adhd_clinical <- factor(d_try$adhd_clinical, levels = c(0, 1), labels = c("no", "yes"))

d_try$adhd_clinical <- as.factor(d_try$adhd_clinical)
d_try <- na.omit(d_try)
ggplot() +
  geom_point(data = d_try, aes(x = days_since_semester_start, y = matrices_hit_rate, color = adhd_clinical), alpha = 0.5) +
  geom_smooth(data = d_try, aes(x = days_since_semester_start, y = matrices_hit_rate, color = adhd_clinical), method = "lm") +     
  labs(
    title = "Investigation of Potential Interaction",
    x = "Days Since Semester Start When Completed the Survey",
    y = "Matrices Test Composite Score",
    caption = "Data source: Data_PostSurvey"
  ) 

```


### Visualization 
Have visualization for conscientioussness, openness, agreeableness, extraversion, matrices hit rate from two datasets; adhd_diagnosis from one dataset 
```{r}
plot_big5_con_post <- ggplot(d_try, aes(x = days_since_semester_start, y = conscientiousness_score)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = TRUE)+
  theme_bw(base_size = 14) + 
  labs(y = 'Big-5 Conscientiousness Postsurvey', 
       x = 'Days since semester starts when completed the survey',
       title = 'Whether Conscientiousness affect \n time to research participation',
       caption = "Data source: Data_PostSurvey") +
  ylim(1, 5) +
  xlim(0, 110) +
  stat_cor(color = "red", size = 6) 

plot_big5_con_post
```

```{r}
ggsave(plot = plot_big5_con_post, 
       filename = file.path(figure_path, 'plot_conscientiousness_postsurvey.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```



```{r}

plot_conscientiousness_replicat <- ggplot(d_personality, aes(x = days_since_semester_start, y = conscientiousness)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(y = 'Big-5 Conscientiousness Score Replication', 
       x = 'Days since semester starts when completed the survey', 
       title = 'Whether Conscientiousness affect \n time to research participation',
       caption = "Data source: Data_Personality") +
  xlim(60, 110) +
  ylim(18, 45) +
  stat_cor(color = "blue", size = 6) 

plot_conscientiousness_replicat

ggsave(plot = plot_conscientiousness_replicat, 
       filename = file.path(figure_path, 'plot_conscientiousness_replicat.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```


```{r}
plot_big5_open_post <- ggplot(d_try, aes(x = days_since_semester_start, y = openness_score)) +
  geom_jitter(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(y = 'Big-5 Openness Postsurvey', 
       x = 'Days since semester starts when completed the survey',
       title = 'Whether Openness affect \n time to research participation',
       caption = "Data source: Data_PostSurvey") +
  xlim(0,110)+
  ylim(1, 5) +
  stat_cor(color = "red", size = 6)
plot_big5_open_post

ggsave(plot = plot_big5_open_post, 
       filename = file.path(figure_path, 'plot_big5_open_post.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```

```{r}
plot_openness_replication <- ggplot(d_personality, aes(x = days_since_semester_start, y = openness)) +
  geom_jitter() +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(y = 'Big-5 Openness Score Replication', 
       x = 'Days since semester starts when completed the survey', 
       title = 'Whether Openness affect \n time to research participation',
       caption = "Data source: Data_Personality") +
  xlim(60, 110) +
  ylim(14, 50) +
  stat_cor(color = "blue", size = 6) 
plot_openness_replication 
ggsave(plot = plot_openness_replication, 
       filename = file.path(figure_path, 'plot_openness_replication.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```


Extraversion
```{r}
plot_big5_extraversion_post <- ggplot(d_try, aes(x = days_since_semester_start, y = extraversion_score)) +
  geom_jitter(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(y = 'Big-5 Extraversion Postsurvey', 
       x = 'Days since semester starts when completed the survey',
       title = 'Whether Extraversion affect \n time to research participation',
      caption = "Data source: Data_PostSurvey") +
  xlim(0,110)+
  ylim(1, 5.4) +
  stat_cor(color = 'red', size = 6)
plot_big5_extraversion_post

ggsave(plot = plot_big5_extraversion_post, 
       filename = file.path(figure_path, 'plot_big5_extraversion_post.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```


```{r}
plot_extraversion_replication <- ggplot(d_personality, aes(x = days_since_semester_start, y = extraversion)) +
  geom_jitter() +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(y = 'Big-5 Extraversion Replication', 
       x = 'Days since semester starts when completed the survey', 
       title = 'Whether Extraversion affect \n time to research participation',
       caption = "Data source: Data_Personality") +
  xlim(60, 110) +
  ylim(15, 36) +
  stat_cor(color = "blue", size = 6) 
plot_extraversion_replication

ggsave(plot = plot_extraversion_replication, 
       filename = file.path(figure_path, 'plot_extraversion_replication.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```

Agreeableness
```{r}
plot_big5_agr_complete <- ggplot(d_try, aes(x = days_since_semester_start, y = agreeableness_score)) +
  geom_jitter(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(y = 'Big-5 Agreeableness Postsurvey', 
       x = 'Days since semester starts when completed the survey',
       title = "Whether Agreeableness affect \n time to research participation",
       caption = "Data source: Data_PostSurvey") +
  xlim(0, 110) +
  ylim(1, 5.5) +
  stat_cor(color = "red", size = 6) 
plot_big5_agr_complete

ggsave(plot = plot_big5_agr_complete, 
       filename = file.path(figure_path, 'plot_big5_agr_complete.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```

```{r}
plot_agreeableness_replication <- ggplot(d_personality, aes(x = days_since_semester_start, y = agreeableness)) +
  geom_jitter() +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(y = 'Big-5 Agreeableness Replication', 
       x = 'Days since semester starts when completed the survey', 
       title = 'Whether Agreeableness affect \n time to research participation',
       caption = "Data source: Data_Personality") +
  xlim(60, 110) +
  ylim(17, 45) +
  stat_cor(color = "blue", size = 6) 
plot_agreeableness_replication
ggsave(plot = plot_agreeableness_replication, 
       filename = file.path(figure_path, 'plot_agreeableness_replication.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```


Matrices Hit Rate
```{r}
plot_matrices_hit_post <- ggplot(d_try, aes(x = days_since_semester_start, y = matrices_hit_rate)) +
  geom_jitter(alpha = 0.6) +
  geom_smooth(method = "lm", color="red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(y = 'Matrices Hit Rate Postsurvey', 
       x = 'Days since semester starts when completed the survey',
       title = 'Whether Fluid Intelligence affect \n time to research participation',
       caption = "Data source: Data_PostSurvey") +
  xlim(0,110)+
  ylim(0, 1.2) +
  stat_cor(color = "red", size = 6)
plot_matrices_hit_post
ggsave(plot = plot_matrices_hit_post, 
       filename = file.path(figure_path, 'plot_matrices_hit_post.png'), 
       height = 5.5, width = 5.5,
       units = 'in')
```

```{r}
plot_matrices_replication <- ggplot(d_matrices, aes(x = days_since_semester_start, y = matrices_hit_rate)) +
  geom_jitter() +
  geom_smooth(method = "lm", color = "brown", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(y = 'Matrices Hit Rate Replication', 
       x = 'Days since semester starts when completed the survey', 
       title = 'Whether Fluid Intelligence affect \n time to research participation',
       caption = "Data source: Data_Matrices") +
  xlim(0, 110) +
  ylim(0, 1) +
  stat_cor(color = "brown", size = 6) 
plot_matrices_replication
ggsave(plot = plot_matrices_replication, 
       filename = file.path(figure_path, 'plot_matrices_replication.png'), 
       height = 5.5, width = 5.5,
       units = 'in')  
```

```{r}
plot_adhd_postsurvey <- ggplot(d_try, aes(x = days_since_semester_start, y = adhd_total)) +
  geom_jitter(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  theme_bw(base_size = 14) + 
  labs(y = 'Total Score of 6-item ASRS Postsurvey', 
       x = 'Days since semester starts when completed the survey', 
       title = 'Whether ADHD Score affect \n time to research participation',
       caption = "Data source: Data_PostSurvey") +
  xlim(0, 110) +
  ylim(0, 7.0) +
  stat_cor(color = "red", size = 6) 
plot_adhd_postsurvey

ggsave(plot = plot_adhd_postsurvey, 
       filename = file.path(figure_path, 'plot_adhd_postsurvey.png'), 
       height = 5.5, width = 5.5,
       units = 'in') 
```

